knitr document van Steensel lab

# Timeseries on TRIP clonal lines
# Introduction

We would like to know how DNA repair processes are influenced by the chromatin
environment in the context double-strand brakes introduced by CRISPR/Cas9.
From time-series data we can calculate kinetic rates of brake repair.
Tao has previously worked out the kinetics in the endogenous locus of the LBR2 gene.



```{r}
library(ggplot2)
library(data.table)
library(reshape2)
library(plyr)
library(rtracklayer)
library("plyr")
library(deSolve)
library(minpack.lm)
library(FME)
library(truncnorm)
library(colorspace)
library(ggbeeswarm)
library(stringr)
library(gridExtra)

spikein = 'GGTACCCATGATCAAC'

DNA_color = c('MMEJ'="#008296", 'NHEJ'="#e63227", '-7'="#008296", '1'="#e63227",
              'wt'="#7f8185", 'other'="#5C5E42", 'broken'="#BBABB1",
              'SSTR'='#00a23a')

get_combinations <- function(domain_vec){
    n_domain = length(domain_vec)
    get_ivec <-function (last_x, n_domain){
    	if (last_x < n_domain){
    		i_vec = c(0, c((last_x+1):n_domain))
    	} else{
    		i_vec = c(0)
    	}
    }

    combine <- function(x_new, x_vec, n_domain){
        new_vec = c(x_vec, x_new)
        if (length(new_vec)==n_domain){
            return(new_vec)
        } else {
            i_vec = get_ivec(max(new_vec), n_domain)
            x_list = lapply(i_vec, combine, x_vec=new_vec,
                            n_domain=n_domain)
            return(do.call(rbind, x_list))
        }
    }

    comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)

    comb = do.call(rbind, comb_list)
    comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))

    name_list = apply(comb_df, 1, function(x){
    	name = paste(domain_vec[x], collapse='-')
    	if (name == ''){
    		name = 'iDomain'
    	}
    	return(name)
    })

    opt_list = lapply(domain_vec, function(d){
    	 apply(comb_df, 1, function(x){
    		opt = domain_vec[x]
    		return(d%in%opt)
    	})
    })

    opt_df = do.call(cbind, opt_list)
    colnames(opt_df) = domain_vec
    rownames(opt_df) = name_list

    return(opt_df)
}

```
```{r}


clone5 = fread('/DATA/projects/DSBrepair/data/cl20190220_chip_clone5/cl20190222_clone5_LAD.txt',
               stringsAsFactors=F, key='barcode',
               col.names=c('seqnames', 'start', 'end', 'barcode', 'score',
                           'strand', 'LAD'))

setkey(clone5, 'barcode')

clone5 = clone5[, c('barcode', 'LAD')]

clone5 = rbind(clone5, list(c('CATTTCTGATCAATAA', 'ATACTATATTTAACGG'), c(1, 0)))



near_dir = '/DATA/projects/DSBrepair/data/tracks_20190711/nearest/'

file_list = list.files(near_dir, pattern='clone5', full.names=T)

file.dt = data.table(str_match(file_list, '.*/clone5_(.*).txt'))
names(file.dt) = c('file', 'target')
domains.dt = file.dt[!is.na(file),
                 fread(file, col.names=c('barcode', 'domain', 'distance')),
				 by=c('target')]

domains.dt[target=='repliseq', target:='late_replicating']

domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domain_cast = data.table::dcast(domains.dt[target%in%domain_levels, ],
                                barcode ~ target, value.var='domain')
setkey(domain_cast, barcode)

domain_cast[domain_cast=='iDomain'] <- 0
domain_cast[domain_cast=='Domain'] <- 1

domain_cast[domain_cast=='iAD'] <- 0
domain_cast[domain_cast=='AD'] <- 1
domain_cast[, (domain_levels) := lapply(.SD, as.numeric), .SDcols = domain_levels]

opt_df = get_combinations(domain_levels)

group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domain_levels[opt]
    out_vec = domain_levels[!opt]
    in_sum = rowSums(domain_cast[,..in_vec])
    out_sum = rowSums(domain_cast[,..out_vec])
    if (length(in_vec)==0){
        in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
        in_group = which(in_sum==length(in_vec))
    } else  {
        in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
})

names(group_list) = rownames(opt_df)
for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    domain_cast[i_vec, group := opt_name]
}


meta.dt = fread('/DATA/projects/DSBrepair/config/cl20190816_metadata_timeseries_LBR2.tsv',
                stringsAsFactors=F, key='ID')

# meta.dt[,ID:=gsub('-', 'no', ID)]
# meta.dt[,guide:=gsub('-', 'no', guide)]

indir = '/DATA/projects/DSBrepair/data/cl20200107_timeseries_LBR2/'


mutations.dt = meta.dt[, fread(paste0(indir, 'indelPCR_counts/', ID, '.count.table'),
                               stringsAsFactors=F,
                               colClasses=c(rep('character', 3), 'integer'),
                               col.names=c('barcode', 'type', 'indel',
                                           'count')), by=ID]
mutations.dt[type=='ssODN', indel:='SSTR']

## barcode integrated at 2 sites
mutations.dt = mutations.dt[!barcode%in%c('TTGCTCTTCCGCAGAT', 'GGTACCAACTCCAGCA'), ]


mutation_merge.dt = merge(mutations.dt, meta.dt[PCR_type=='indelPCR', ], by='ID')

bc_data = data.table(barcode=clone5$barcode, clone=5)
bc_data[barcode=='CTCTTAATCGCTGCC', barcode:='GCTCTTAATCGCTGCC']

mut_selection = merge(mutation_merge.dt, bc_data, by=c('barcode', 'clone'))

key_vec = c('barcode', 'clone', 'indel', 'series',
            'plasmid', 'shield', 'time')

indel.dt = mut_selection[, list(count=sum(count)),
                                by=key_vec]


sum_count.dt =  indel.dt[series%in%c('TimeSeries2_a', 'TimeSeries2-3_b') &
                         plasmid=='LBR2',
                         list(count=sum(count[indel%in%c('1','-7')])),
                              by=c('barcode', 'clone', 'series',
                                        'plasmid', 'shield', 'time')]
sum_count.dt[,thresh:=all(count>20),
             by=c('barcode', 'clone',
                  'plasmid', 'shield', 'time')]

pdf('cl20200309_indel_count.pdf', useDingbats=F, height=10, width=15)
ggplot(sum_count.dt, aes(x=time, y=log10(count), color=series, shape=thresh)) +
    geom_point() +
    facet_wrap(~barcode)

dev.off()


sum_count_SSTR.dt =  mut_selection[series%in%c('TimeSeries3_a', 'TimeSeries4_a') &
                                   plasmid=='LBR2',
                                   list(count=sum(count[indel%in%c('SSTR','-7')])),
                                        by=c('barcode', 'clone', 'series',
                                             'plasmid', 'shield', 'time')]

sum_count_SSTR.dt[,thresh:=all(count>20),
                  by=c('barcode', 'clone',
                       'plasmid', 'shield', 'time')]

pdf('cl20200309_indel_count_SSTR.pdf', useDingbats=F, height=10, width=15)
ggplot(sum_count_SSTR.dt, aes(x=time, y=log10(count), color=series, shape=thresh)) +
    geom_point() +
    facet_wrap(~barcode)

dev.off()


norm_count.dt = mut_selection[series%in%c('TimeSeries2_a', 'TimeSeries2-3_b') &
                              plasmid=='LBR2',
                              list(barcode=barcode,
                                   indel=indel,
                                   norm=count/sum(count)),
                              by=c('clone', 'series',
                                   'plasmid', 'shield', 'time')]



mean_count.dt = norm_count.dt[, list(mean_norm=mean(norm)),
                              by=c('clone', 'barcode', 'indel',
                                   'plasmid', 'shield', 'time')]


ratio.dt = mean_count.dt[indel%in%c('1', '-7'),
                         list(MMEJ_MMEJNHEJ=mean_norm[indel=='-7']/sum(mean_norm)),
                         by=c('clone', 'barcode', 'plasmid', 'shield', 'time')]




ratio_merge = merge(ratio.dt, unique(sum_count.dt[, c('clone', 'plasmid',
                                                     'barcode', 'shield',
                                                     'time', 'thresh')]),
                   by=c('barcode', 'clone', 'plasmid', 'shield', 'time'))



model_dt <- function(dt, yname){
    y = unlist(dt[,..yname,with=T])
    x = dt$time
    model = 'y~a * exp(-b * exp(-c * x))'
    start_list = list(a=20, b=10, c=0.1)
    test = tryCatch(nls(formula(model),
                        start=start_list,
                        data=data.frame(x,y)),
                    error=function(x){return(NULL)})
    if (!is.null(test)){
        return(as.list(c(coef(test), rse=summary(test)$sigma)))
    }
}

ratio_merge[, group:=domain_cast[barcode,'group']]

gompertz.dt = ratio_merge[thresh==T, model_dt(.SD, 'MMEJ_MMEJNHEJ'),
                          by=c('barcode', 'clone', 'plasmid', 'shield')]



t_gomp = seq(0,ratio_merge[,max(time)], by=0.1)


gompertz_fit.dt = gompertz.dt[, list(time=t_gomp,
                                     MMEJ_MMEJNHEJ = a * exp(-b * exp(-c * t_gomp))),
                              by=c('barcode', 'clone', 'plasmid', 'shield')]


pdf('cl20200420_gompertz_MMEJ_MMEJNHEJ.pdf', useDingbats=F)
ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)

dev.off()



pdf('cl20200420_gompertz_MMEJ_MMEJNHEJ_split.pdf', useDingbats=F, width=15,
    height=10)
ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1) +
    facet_wrap(~barcode)

dev.off()


```



```{r}


ratio.dt = indel.dt[series%in%c('TimeSeries2_a', 'TimeSeries2-3_b') &
                         plasmid=='LBR2' & indel%in%c('1', '-7'),
                         list(MMEJ_MMEJNHEJ=count[indel=='-7']/sum(count)),
                         by=c('clone', 'barcode', 'plasmid', 'series', 'shield',
                              'time')]

mean_ratio.dt = ratio.dt[, list(MMEJ_MMEJNHEJ=mean(MMEJ_MMEJNHEJ)),
                         by=c('clone', 'barcode', 'plasmid', 'shield',
                              'time')]


ratio_merge = merge(mean_ratio.dt, unique(sum_count.dt[, c('clone', 'plasmid',
                                                           'barcode', 'shield',
                                                           'time', 'thresh')]),
                   by=c('barcode', 'clone', 'plasmid', 'shield', 'time'))

ratio_merge[, group:=domain_cast[barcode,'group']]

model_dt <- function(dt, yname){
    y = unlist(dt[,..yname,with=T])
    x = dt$time
    model = 'y~a * exp(-b * exp(-c * x))'
    start_list = list(a=20, b=10, c=0.1)
    test = tryCatch(nls(formula(model),
                        start=start_list,
                        data=data.frame(x,y)),
                    error=function(x){return(NULL)})
    if (!is.null(test)){
        return(as.list(c(coef(test), rse=summary(test)$sigma)))
    }
}

gompertz.dt = ratio_merge[thresh==T, model_dt(.SD, 'MMEJ_MMEJNHEJ'),
                          by=c('barcode', 'clone', 'plasmid', 'shield', 'group')]

t_gomp = seq(0,ratio_merge[,max(time)], by=0.1)


gompertz_fit.dt = gompertz.dt[, list(time=t_gomp,
                                     MMEJ_MMEJNHEJ = a * exp(-b * exp(-c * t_gomp))),
                              by=c('barcode', 'clone', 'plasmid', 'shield', 'group')]



pdf('cl20200420_gompertz_MMEJ_MMEJNHEJ_no_norm.pdf', useDingbats=F, height=5)
ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)

ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ]) +
    theme_bw() +
    theme(aspect.ratio=1)


ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)


ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ], size=0.5) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)

dev.off()



pdf('cl20200420_gompertz_MMEJ_MMEJNHEJ_split_no_norm.pdf', useDingbats=F, width=15,
    height=10)
ggplot(gompertz_fit.dt, aes(x=time, y=MMEJ_MMEJNHEJ, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1) +
    facet_wrap(~barcode)

dev.off()


title_vec = c('a'='a: asymptote (top of the slope)',
              'b'='b: displacement along x-axis (delay)',
              'c'='c: growth rate')

gompertz.dt[!group%in%c('iDomain', 'late_replicating-LAD-H3K9me2',
                             'H3K27me3'), group:='other']

pdf('cl20200420_gompertz_MMEJ_MMEJNHEJ_variables.pdf', useDingbats=F, width=3.5,
    height=10)

plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz.dt, aes_string(x='group', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(title_vec[x]) +
        theme_bw()
    })

do.call(grid.arrange, plot_list)

plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz.dt, aes_string(x='group', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(title_vec[x]) +
        theme_bw() +
        expand_limits(y=0)
    })

do.call(grid.arrange, plot_list)


plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz.dt, aes_string(x='1', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(x) +
        theme_bw() +
        ggtitle(title_vec[x]) +
        expand_limits(y=0)
    })

do.call(grid.arrange, plot_list)

dev.off()

```



```{r}


ratio_SSTR.dt = indel.dt[series%in%c('TimeSeries3_a', 'TimeSeries4_a') &
                         plasmid=='LBR2' & indel%in%c('SSTR', '-7'),
                         list(MMEJ_MMEJSSTR=count[indel=='-7']/sum(count)),
                         by=c('clone', 'barcode', 'plasmid', 'series', 'shield',
                              'time')]

mean_ratio_SSTR.dt = ratio_SSTR.dt[, list(MMEJ_MMEJSSTR=mean(MMEJ_MMEJSSTR)),
                                   by=c('clone', 'barcode', 'plasmid', 'shield',
                                   'time')]


ratio_merge_SSTR = merge(mean_ratio_SSTR.dt,
                         unique(sum_count_SSTR.dt[, c('clone', 'plasmid',
                                                      'barcode', 'shield',
                                                      'time', 'thresh')]),
                         by=c('barcode', 'clone', 'plasmid', 'shield', 'time'))

ratio_merge_SSTR[, group:=domain_cast[barcode,'group']]


gompertz_SSTR.dt = ratio_merge_SSTR[thresh==T, model_dt(.SD, 'MMEJ_MMEJSSTR'),
                                    by=c('barcode', 'clone', 'plasmid',
                                         'shield', 'group')]

t_gomp = seq(0,ratio_merge[,max(time)], by=0.1)


gompertz_fit_SSTR.dt = gompertz_SSTR.dt[, list(time=t_gomp,
                                        MMEJ_MMEJSSTR = a * exp(-b * exp(-c * t_gomp))),
                                        by=c('barcode', 'clone', 'plasmid',
                                             'shield', 'group')]



pdf('cl20200420_gompertz_MMEJ_MMEJSSTR_no_norm.pdf', useDingbats=F, height=5)
ggplot(gompertz_fit_SSTR.dt, aes(x=time, y=MMEJ_MMEJSSTR, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge_SSTR[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)

ggplot(gompertz_fit_SSTR.dt, aes(x=time, y=MMEJ_MMEJSSTR, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge_SSTR[thresh==T, ]) +
    theme_bw() +
    theme(aspect.ratio=1)


ggplot(gompertz_fit_SSTR.dt, aes(x=time, y=MMEJ_MMEJSSTR, group=barcode,
                            color=group)) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)


ggplot(gompertz_fit_SSTR.dt, aes(x=time, y=MMEJ_MMEJSSTR, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge_SSTR[thresh==T, ], size=0.5) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1)

dev.off()



pdf('cl20200420_gompertz_MMEJ_MMEJSSTR_split_no_norm.pdf', useDingbats=F, width=15,
    height=10)
ggplot(gompertz_fit_SSTR.dt, aes(x=time, y=MMEJ_MMEJSSTR, group=barcode,
                            color=group)) +
    geom_point(data=ratio_merge_SSTR[thresh==T, ]) +
    geom_line() +
    theme_bw() +
    theme(aspect.ratio=1) +
    facet_wrap(~barcode)

dev.off()


gompertz_SSTR.dt[!group%in%c('iDomain', 'late_replicating-LAD-H3K9me2',
                             'H3K27me3'), group:='other']

pdf('cl20200420_gompertz_MMEJ_MMEJSSTR_variables.pdf', useDingbats=F, width=3.5,
    height=10)

plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz_SSTR.dt, aes_string(x='group', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(title_vec[x]) +
        theme_bw()
    })

do.call(grid.arrange, plot_list)

plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz_SSTR.dt, aes_string(x='group', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(title_vec[x]) +
        theme_bw() +
        expand_limits(y=0)
    })

do.call(grid.arrange, plot_list)


plot_list = lapply(c('a','b','c'), function(x){
    ggplot(gompertz_SSTR.dt, aes_string(x='1', y=x, color='group')) +
        geom_quasirandom() +
        ggtitle(x) +
        theme_bw() +
        ggtitle(title_vec[x]) +
        expand_limits(y=0)
    })

do.call(grid.arrange, plot_list)

dev.off()


saveRDS(domain_cast, '../data/R/cl20200420_domains_clone5.RDS')
```
