---
title: "Figure 6 and supplemental 6"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r}
StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

## Select outdir
out.dir = paste0("figures/rs", Date, "/")
dir.create(out.dir)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# INPUT DIRECTORY
in.dir.date = 20201029
in.dir = paste0("/DATA/projects/DSBrepair/data/R/rs", in.dir.date, "/")

# libraries:
library(ggplot2)
library(ggpubr)
# library(reshape2)
# library(tibble)
# library(GenomicRanges)
# library(rtracklayer)
library(corrr)
library(Hmisc)
library(ggbeeswarm)
library(RColorBrewer)
library(data.table)
library(dplyr)
# library(plyr)
library(tidyr)
# library(stringr)
# library(plotly)
library(ggpmisc)
library(glmnet)
library(cowplot)
library(mda)
library(earth)
library(yaml)
library(vip)
library(caret)
library(scales)
library(glmnet)
library(gridExtra)
library(ggcorrplot)
library(bnlearn)
library(pheatmap)
library(ppcor)
library(parallel)
library(stringr)

color_redblue <- rev(brewer.pal(11,"RdBu"))
duocolor <- c("#EE756E", "#2BB6BE")
colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colores <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colori  <- c("other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "ssODN" = "#007A4C")
colora <- c("0" = "#808184", "1" = "#E1251B", "-7" = "#223DA0", "-14" = "#223DA0")
coloru <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "MMEJplusNHEJplusHDR" = "#EE8A2A", "HDR" = "#007A4C", "MMEJplusHDR" = "#EE8A2A")
colory <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "SSTR" = "#007A4C")

measure_color = c(NHEJ='#e1251b', MMEJ='#26419a', SSTR='#007A4C', None='grey')

method_color = c(MARS='#00bfc4', LASSO='#f8766d')
lad_color <- c(LAD="#802cdc", iLAD="#dfaf2f")
sign_color <- c(down="#AA6375", up="#7B9A36")

feature_fill = c(insulator="#a6b5a3", repressive="#304441", euchromatin="#cda4cc",
                 transcribing="#ED9114", HDAC="#aac9e0", accessibility='#cebc85',
                 methylation='#7dc98f')

feature_color = c(insulator="black", repressive="white", euchromatin="black",
                  transcribing="black", HDAC="black", accessibility='black',
                  methylation='black')

clustering = data.frame(row.names=c('LMNB1', 'late_replicating', 'H3K9me2',
                                    'H3K27me3', 'EZH2', 'CTCF', 'SMC3',
                                    'HDAC3', 'HDAC2', 'HDAC1', 'H3K4me1',
                                    'H3K4me2', 'H3K4me3', 'H3K27ac',
                                    'H4K5acK8ac', 'H2AFZ', 'DNAse', 'Dam', 'm5C',
                                    'H3K79me2', 'TTseq', 'H3K36me3', 'POL2AS2',
                                    'POL2'),
                        group=factor(c(rep('repressive', 5), rep('insulator', 2),
                                       rep('HDAC', 3), rep('euchromatin', 6),
                                       rep('accessibility', 2), 'methylation',
                                       rep('transcribing', 5)),
                                     levels=c('transcribing', 'accessibility',
                                              'methylation', 'euchromatin',
                                              'HDAC', 'insulator', 'repressive')))

clone5barcodes <- c("AGGGCGTAAAATATTT.B",
                    "TATGGCTGTCGGGTAG.B",
                    "TGTCCCTTAGTACTTT.B",
                    "AGAAAATAATATGACG.B",
                    "CGGCCTGAAGGTCAGG.B",
                    "TTGAACGCGGGCTCGG.B",
                    "GCTAACATCACGAATC.B",
                    "GCGCACCCTTTAATTG.B",
                    "ACTGTCGAGTTGTCCG.B",
                    "CCGGGGACGTATGCAC.B",
                    "TCTTTTGAGGAGCTGA.B",
                    "ATATCGTTGCTGGAGA.B",
                    "CATCCACCACACTTCA.B",
                    "ACCCCTAAAGGCGCTG.B",
                    "ATACTATATTTAACGG.B",
                    "CATTTCTGATCAATAA.B",
                    "GAGCGCGTCACCGGGT.B",
                    "GTACCTCTCGATAGTG.B",
                    "TGGCCAATATTTGTCT.B")

```

### Custom functions
Functions used thoughout this script.
```{r functions, message=FALSE, warnings=FALSE}
cor_test <- function(y, X, left, right){
  cor_list = lapply(X, function(x){
    cor.test(x,y)
  })
  p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
  cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
  p_adj = p.adjust(p_vec)
  color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                 'None')
  importance = abs(cor_vec)
  target=factor(colnames(X), levels=rownames(clustering))
  feature_group = clustering[colnames(X), ]
  return(list(correlation=cor_vec,importance = importance, p=p_vec,
              p_adj=p_adj, target=target, color=color,
              feature_group=feature_group))
}


get_combinations <- function(domain_vec){
  n_domain = length(domain_vec)
  get_ivec <-function (last_x, n_domain){
    if (last_x < n_domain){
      i_vec = c(0, c((last_x+1):n_domain))
    } else{
      i_vec = c(0)
    }
  }
  
  combine <- function(x_new, x_vec, n_domain){
    new_vec = c(x_vec, x_new)
    if (length(new_vec)==n_domain){
      return(new_vec)
    } else {
      i_vec = get_ivec(max(new_vec), n_domain)
      x_list = lapply(i_vec, combine, x_vec=new_vec,
                      n_domain=n_domain)
      return(do.call(rbind, x_list))
    }
  }
  
  comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)
  
  comb = do.call(rbind, comb_list)
  comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))
  
  name_list = apply(comb_df, 1, function(x){
    name = paste(domain_vec[x], collapse='-')
    if (name == ''){
      name = 'iDomain'
    }
    return(name)
  })
  
  opt_list = lapply(domain_vec, function(d){
    apply(comb_df, 1, function(x){
      opt = domain_vec[x]
      return(d%in%opt)
    })
  })
  
  opt_df = do.call(cbind, opt_list)
  colnames(opt_df) = domain_vec
  rownames(opt_df) = name_list
  
  return(opt_df)
}

test_wilcox <- function(dt, group_by='condition', y_name='MMEJ_MMEJNHEJ'){
  bc_vec = dt[,unique(barcode)]
  name1 = unlist(unique(dt[,group_by, with=F])[1])
  dt_c = dt[get(group_by)==eval(name1), ]
  dt_s = dt[get(group_by)!=eval(name1), ]
  setkey(dt_c, 'barcode')
  setkey(dt_s, 'barcode')
  
  w = wilcox.test(unlist(dt_c[bc_vec,..y_name]),
                  unlist(dt_s[bc_vec,..y_name]),
                  paired=T)
  return(list(p_value = w$p.value))
}



plot_comb_grid_beeswarm <- function(dt, opt_df, domains, y_name, min_count=0,
                                    max_count=Inf, group_by=NULL, color=NULL, y_plabel=c(-1.5, 1.5)){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(dt[,..in_vec])
    out_sum = rowSums(dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(dt[,table(group)])
  } else {
    group_count_df = data.frame(dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]
  
  # wilcox_list = lapply(group_levels[group_levels!='iDomain'],
  #                      function(g){
  #                          x = indel_selection[group==g, y_name, with=F]
  #                          y = indel_selection[group=='iDomain', y_name,
  #                                              with=F]
  #                          wilcox.test(unlist(x),unlist(y))
  #                      })
  
  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    theme_bw() +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color=FALSE)  +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", " "))
  
  
  if (is.null(group_by)){
    beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=0.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_compare_means(aes(label = ..p.signif..),
                         symnum.args = symnum,
                         method = 'wilcox.test',
                         ref.group='iDomain')
    
  } else {
    wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                by='group']
    wilcox_dt[,p_adj:=p.adjust(p_value)]
    wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                symbols=symnum$symbols)]
    
    
    beeswarm = ggplot(indel_selection,
                      aes_string(x='group', y=y_name, color=group_by)) +
      geom_quasirandom(dodge.width=1) +
      scale_color_manual(values=color) +
      stat_summary(aes_string(group=group_by), fun=median,
                   fun.min = median, fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red',
                   position= position_dodge(width =1)) +
      geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                inherit.aes=F)
    print(wilcox_dt)
  }
  beeswarm = beeswarm +
    theme_bw() +
    geom_vline(xintercept=seq(1.5, length(unique(dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank()) +
    ylim(y_plabel)
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme_bw() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, beeswarm, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}


PlotDataTracks <- function(bins, chr = "chr1", start = 1, end = 1e6, samples = NA, ipr_loc = 1, cols = NA) {
  
  # Get the data
  data <- bins[seqnames(bins) == chr & 
                 start(bins) >= start & 
                 end(bins) <= end,]
  
  data <- as(data, "data.frame")
  
  if (any(!is.na(samples))) {
    samplecols = which(colnames(data) %in% samples)
    data <- data[, c(1:5, samplecols)]
  } 
  
  # Melt the data
  data.melt <- reshape2::melt(data, id.vars = c("seqnames", "start", "end", "width", "strand"))
  data.melt <- data.melt[complete.cases(data.melt), ]
  
  # Plot the data
  plt <- ggplot(data.melt, aes(xmin = start / 1e6, xmax = end / 1e6, ymin = 0, ymax = value,
                               fill = variable)) +
    geom_rect() +
    geom_hline(yintercept = 0, size = 0.5) +
    facet_grid(variable ~ .) +
    xlab(paste0(chr, " (Mb)")) +
    ylab("Score") +
    geom_vline(xintercept = ipr_loc / 1e6) +
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0)) +
    scale_fill_manual(values = cols) +
    theme_classic()
  
  plot(plt)
  
}
```

# Data import
```{r import, message=FALSE, warnings=FALSE}
trip_tib_mut_2000 = readRDS(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RDS"))
trip_tib_2000 = readRDS(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RDS"))
load(paste0(in.dir, "Analyis_Mapping_RSTP2_2000.RData"))


# load(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RData"))
# load(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RData"))


clone5_domains <- readRDS("/DATA/projects/DSBrepair/data/R/rs20200311_domains_clone5.RDS")

ipr_order = c("IPR16", "IPR6", "IPR14", "IPR7", "IPR17", "IPR13", 
              "IPR8", "IPR9", "IPR12", "IPR11", "IPR2", "IPR10", 
              "IPR1", "IPR18", "IPR4", "IPR15", "IPR5", "IPR19", 
              "IPR3")

clone5_domains[,IPR:=factor(IPR, levels=ipr_order)]
```

```{r import pAdamID data, message=FALSE, warnings=FALSE}
# Load dependencies
suppressMessages(suppressWarnings(library(GenomicRanges)))
suppressMessages(suppressWarnings(library(rtracklayer)))
#

# Load the data sets required
file_names <- c(#"LMNA_KO", 
  "LMNA_KO1", "LMNA_KO2", "LMNA_KO3","LMNA_KO4",
  #"LBR_KO", 
  "LBR_KO1", "LBR_KO2", "LBR_KO3", "LBR_KO4",
  "Clone5")

# pAdam_dir = '/DATA/usr/t.v.schaik/proj/lamin_KO/ts200329_Ruben_clones_pADamID/results/tracks/normalized/bin-20kb/'
# file_list = list.files(pAdam_dir, pattern='combined.bw', full.names=T)


pAdam_dir = '/DATA/projects/DSBrepair/data/tracks_20200415/track/'
file_list = list.files(pAdam_dir, pattern='20kb.bw', full.names=T)

# LMNA_KO = import(file_list[])
LMNA_KO1 = import(file_list[8]) 
LMNA_KO2 = import(file_list[9])
LMNA_KO3 = import(file_list[10])
LMNA_KO4 = import(file_list[11])

# LBR_KO = import(file_list[5])
LBR_KO1 = import(file_list[4])
LBR_KO2 = import(file_list[5])
LBR_KO3 = import(file_list[6])
LBR_KO4 = import(file_list[7])

Clone5 = import(file_list[12])

data_tracks <- list(#LMNA_KO, 
  LMNA_KO1, LMNA_KO2,LMNA_KO3, LMNA_KO4,
  #LBR_KO, 
  LBR_KO1, LBR_KO2, LBR_KO3, LBR_KO4,
  Clone5)
names(data_tracks) <- file_names

# Let's combine the various bigwig files into one
# First, create a GRanges with all the bins
bins <- read.table("/DATA/usr/t.v.schaik/proj/tests/results/ts180813_GCF5083_pADamIDtests/results/counts/bin-20kb/pADamID-Hap1_r4_Dam_semibad-Dam-20kb.counts.txt.gz",
                   sep = "\t")[, 1:3]
names(bins) <- c("seqnames", "start", "end")
bins$start <- bins$start + 1
bins <- as(bins, "GRanges")

# Add the data
mcols(bins)[, file_names] <- NA 

for (i in file_names) {
  track <- data_tracks[[i]]
  ovl <- findOverlaps(track, bins)
  mcols(bins)[subjectHits(ovl), i] <- (track$score[queryHits(ovl)] - mean(track$score)) / sd(track$score)
}


# add mean of each KO for plotting
bins$LMNA_KO = rowMeans(as(mcols(bins)[, c("LMNA_KO1", "LMNA_KO2", "LMNA_KO3", "LMNA_KO4")],"data.frame"), na.rm = T)
bins$LBR_KO = rowMeans(as(mcols(bins)[, c("LBR_KO1", "LBR_KO2", "LBR_KO3", "LBR_KO4")],"data.frame"), na.rm = T)


```

```{r christ damid data processing}
dam_list = list.files('/DATA/projects/DSBrepair/data/tracks_20200415/dam/',
                      pattern='clone_set-.*LMNB2-', full.names=T)

file_dt = data.table(str_match(dam_list, '.*clone_set-([0-9]+).*-(.*).txt'))

colnames(file_dt) = c('file', 'binsize', 'KO')
file_dt[,binsize:=as.numeric(binsize)]

dam_dt = file_dt[, fread(file,stringsAsFactors=F), by=c('binsize', 'KO')]


# bw_list = lapply(dam_dt[binsize!=2000,unique(binsize)], function(bs){
#         dir = paste0('/DATA/usr/t.v.schaik/proj/lamin_KO/',
#                      'ts200329_Ruben_clones_pADamID/results/',
#                      'tracks/normalized/bin-', as.numeric(bs)/1000, 'kb/')
#         list.files(dir, pattern='combined', full.names=T)
#     })
# 
# bw_dt = data.table(str_match(unlist(bw_list), '.*bin-([0-9]+)kb.*_5_(.*)_LMNB2.*'))
# colnames(bw_dt) = c('file', 'kb', 'KO')
# bw_dt[, binsize := 1000 * as.numeric(kb)]
# 
# locations = fread('/DATA/projects/DSBrepair/data/cl20200421_mapped_barcodes_clones.tsv',
#                   stringsAsFactors=F,
#                   col.names=c('seqnames', 'end', 'barcode', 'clone'))
# locations[,start:=end-1]
# 
# bc_gr = makeGRangesFromDataFrame(locations, keep.extra.columns=T)
# 
# find_overlaps <- function(bw_file, bc_gr){
#     bw_gr=import.bw(bw_file)
#     o = distanceToNearest(bc_gr, bw_gr)
#     return(data.table(barcode=bc_gr$barcode,
#                       score=score(bw_gr[subjectHits(o)]),
#                       dist=mcols(o)$distance))
# }
# 
# track_score = bw_dt[, find_overlaps(file, bc_gr), by=c('binsize', 'KO')]
# 
# 
# dam_merge = merge(dam_dt, track_score, by.x=c('name', 'binsize', 'KO'),
#                   by.y=c('barcode', 'binsize', 'KO'))
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>%
  filter(exp_pool  %in% c("mean_RSTP2_2000_LBR2_64",
                          "mean_RSTP2_2000_LBR2_BIX01294_64",
                          "mean_RSTP2_2000_LBR2_DMSOGSK_64",
                          "mean_RSTP2_2000_LBR2_GSK126_64"))

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_64",
                         "mean_RSTP2_2000_LBR2_BIX01294_64",
                         "mean_RSTP2_2000_LBR2_DMSOGSK_64",
                         "mean_RSTP2_2000_LBR2_GSK126_64"))

mean_trip_tib_2000 <- trip_tib_2000 %>%
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```
# Figures
## Main Figures

### Panel A
```{r Fig6A, fig.height=5, fig.width=7}
domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

indel.dt = trip_tib %>% filter(drug %in% c("DMSO_GSK", "GSK126"), cell_line == "RSTP2_2000") %>%
  dplyr::select(barcode, drug, MMEJ_MMEJNHEJ, domains) %>%
  spread(drug, MMEJ_MMEJNHEJ) %>%
  filter(!is.na(GSK126), !is.na(DMSO_GSK)) %>%
  mutate(GSK_ratio = log2(GSK126/DMSO_GSK)) %>%
  distinct(barcode, .keep_all = TRUE) %>%
  data.table()

dim(indel.dt)

indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)

# Ratio of SSTR vs MME
plot_comb_grid_beeswarm(indel.dt, opt_df,
                        domain_levels, 'GSK_ratio', min_count = 1, y_plabel=c(-1.5, 1.5))

k27me3 = indel.dt %>% filter(late_replicating == 0, 
                             LAD == 0,
                             H3K9me2 == 0,
                             H3K27me3 == 1) %>% 
  pull(GSK_ratio)

hetcomb = indel.dt %>% filter(late_replicating == 1, 
                              LAD == 1,
                              H3K9me2 == 1,
                              H3K27me3 == 0) %>% 
  pull(GSK_ratio)

idomain = indel.dt %>% filter(late_replicating == 0, 
                              LAD == 0,
                              H3K9me2 == 0,
                              H3K27me3 == 0) %>% 
  pull(GSK_ratio)

wilcox.test(idomain, k27me3)
wilcox.test(idomain, hetcomb)


# barcodes_DMSO <- filter(trip_tib, drug == "DMSO_GSK") %>% pull(barcode) %>% unique()
# barcodes_GSK <- filter(trip_tib, drug == "GSK126") %>% pull(barcode) %>% unique()
# barcodes_both <- intersect(barcodes_DMSO, barcodes_GSK)
# 
# 

# bcstest <- trip_tib %>% filter(drug %in% c("DMSO_GSK", "GSK126"), cell_line == "RSTP2_2000") %>%
#     dplyr::select(barcode, drug, MMEJ_MMEJNHEJ, domains) %>%
#     spread(drug, MMEJ_MMEJNHEJ) %>%
#     filter(!is.na(GSK126), !is.na(DMSO_GSK)) %>%
#     mutate(GSK_ratio = GSK126/DMSO_GSK) %>%
#     distinct(barcode, .keep_all = TRUE) %>% filter(abs(log2(GSK_ratio)) > 1) %>% pull(barcode)
# 
# 
# p <- ggpaired(filter(trip_tib, drug %in% c("DMSO_GSK", "GSK126"), barcode %in% bcstest), x = "drug", y = "MMEJ_MMEJNHEJ", id = "barcode",
#             color = "drug", palette = "jco",
#              line.color = "gray", line.size = 0.4,
#              facet.by = "H3K27me3_domain", short.panel.labs = FALSE)
# 
# # Use only p.format as label. Remove method name.
# p + stat_compare_means(label = "p.format", paired = TRUE)
# 
# 
# pdf('rs20201013_GSKguideCheck.pdf', useDingbats=F, width=8, height=20)
# ggplot(filter(trip_tib_mut, drug %in% c("DMSO_GSK", "GSK126"), barcode %in% bcstest), aes(mutation, freq)) + geom_bar(stat = "identity") + facet_grid(barcode ~ drug)
# dev.off
# 
# 
# 
# 
# indel.dt %>% unite("code", c(late_replicating, LAD, H3K9me2, H3K27me3)) %>% 
#   mutate(group = ifelse(code == "1_1_1_0", "triple", 
#                         ifelse(code == "0_0_0_0", "euchrom", 
#                                ifelse(code == "0_0_0_1", "H3K27", "other")))) %>% 
#   gather(DMSO_GSK, GSK126, key = "drug", value = "MMEJ_MMEJNHEJ") %>% 
#   ggpaired(., x = "drug", y = "MMEJ_MMEJNHEJ", id = "barcode",
#            color = "drug", palette = "jco", 
#            line.color = "gray", line.size = 0.4,
#            facet.by = "group", short.panel.labs = FALSE) + 
#   stat_compare_means(label = "p.format", paired = TRUE)
```


### Panel B
```{r Fig6B, fig.height= 4, fig.width= 8}
trip_tib_KO = trip_tib_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_clone5_LBR2_LBR_KO_1_64", "mean_RSTP2_clone5_LBR2_LBR_KO_2_64", 
                         "mean_RSTP2_clone5_LBR2_LBR_KO_3_64", "mean_RSTP2_clone5_LBR2_LBR_KO_4_64", 
                         "mean_RSTP2_clone5_LBR2_LMNA_KO_1_64", "mean_RSTP2_clone5_LBR2_LMNA_KO_2_64", 
                         "mean_RSTP2_clone5_LBR2_LMNA_KO_3_64", "mean_RSTP2_clone5_LBR2_LMNA_KO_4_64", 
                         "mean_RSTP2_clone5_LBR2_KOCTRL1_64", 
                         "mean_RSTP2_clone5_LBR2_KOCTRL2_64", 
                         "mean_RSTP2_clone5_LBR2_KOCTRL3_64"),
         barcode %in% clone5barcodes) %>%
  distinct(barcode, exp_pool, .keep_all = TRUE) %>% 
  mutate(KO = ifelse(drug %in% c("LMNA_KO_1",
                                 "LMNA_KO_2",
                                 "LMNA_KO_3",
                                 "LMNA_KO_4"), 
                     "LMNA_KO", 
                     ifelse(drug %in% c("LBR_KO_1", 
                                        "LBR_KO_2",
                                        "LBR_KO_3",
                                        "LBR_KO_4"), 
                            "LBR_KO", 
                            "Clone5")))

indel.dt = trip_tib_KO %>%
  dplyr::select(barcode, drug, MMEJ_MMEJNHEJ, domains, LMNB1) %>%
  spread(drug, MMEJ_MMEJNHEJ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(KO_CTRL = mean(KOCTRL1, KOCTRL2, KOCTRL3)) %>% 
  mutate(LMNA_KO1_ratio = LMNA_KO_1/KO_CTRL,
         LMNA_KO2_ratio = LMNA_KO_2/KO_CTRL,
         LMNA_KO3_ratio = LMNA_KO_3/KO_CTRL,
         LMNA_KO4_ratio = LMNA_KO_4/KO_CTRL,
         LBR_KO1_ratio = LBR_KO_1/KO_CTRL,
         LBR_KO2_ratio = LBR_KO_2/KO_CTRL,
         LBR_KO3_ratio = LBR_KO_3/KO_CTRL,
         LBR_KO4_ratio = LBR_KO_4/KO_CTRL) %>%
  distinct(barcode, .keep_all = TRUE) %>%
  left_join(clone5_domains[, c(1,6)], by = "barcode") %>%
  data.table()


ko_group_colors = c("LMNA_KO" = "#f9b016",
                    "LBR_KO" = "#006c68",
                    "Clone5" = "black")

# ggplot(trip_tib_KO, aes(LAD_domain, MMEJ_MMEJNHEJ, fill = KO)) +
#   geom_boxplot() + scale_fill_manual(values = ko_group_colors)

clone5_domains

ko_colors = c("LMNA_KO_1" = "#f9b016",
              "LMNA_KO_2" = "#f9b016",
              "LMNA_KO_3" = "#f9b016",
              "LMNA_KO_4" = "#f9b016",
              "LBR_KO_1" = "#006c68", 
              "LBR_KO_2" = "#006c68",
              "LBR_KO_3" = "#006c68",
              "LBR_KO_4" = "#006c68",
              "KOCTRL1" = "black",
              "KOCTRL2" = "black",
              "KOCTRL3" = "black")

ko_group_colors = c("LMNA_KO" = "#f9b016",
                    "LBR_KO" = "#006c68",
                    "Clone5" = "black")

trip_tib_KO_domains = trip_tib_KO %>% 
  left_join(clone5_domains, by = "barcode") %>%  # add clone5 domains to table
  mutate(KO_clone = gsub('(.*)_KO_([0-9]+)', '\\1\\2', drug)) # make KO_clone colum

trip_tib_KO_domains[trip_tib_KO_domains$group == "iDomain", ]$group <- "euchromatin" 
trip_tib_KO_domains[trip_tib_KO_domains$group == "LAD-H3K9me2", ]$group <- "other"
trip_tib_KO_domains[trip_tib_KO_domains$group == "late_replicating-H3K9me2", ]$group <- "other"
trip_tib_KO_domains[trip_tib_KO_domains$group == "late_replicating", ]$group <- "other"
trip_tib_KO_domains[trip_tib_KO_domains$group == "late_replicating-LAD-H3K9me2", ]$group <- "triple"

group_colors = c("euchromatin" = "#f7941d",
                 "other" = "#838687",
                 "triple" = "#662d91",
                 "H3K27me3" = "#d21f8a")

# 
# ggplot(trip_tib_KO_domains, aes(factor(IPR, levels = ipr_order), MMEJ_MMEJNHEJ)) +
#   geom_point(aes(color = group), size = 2) +
#   scale_color_manual(values = group_colors) +
#   theme_bw(base_size = 16) +
#   facet_grid(~ KO, scales = "free")
# 
# ggplot(trip_tib_KO_domains, aes(factor(IPR, levels = ipr_order), MMEJ_MMEJNHEJ, shape = KO)) + 
#   geom_quasirandom(aes(color = KO), size = 2) +
#   scale_color_manual(values = ko_group_colors) + 
#   theme_bw(base_size = 16)+ 
#   theme(axis.text.x = element_text(angle = 90)) +
#   ylim(0, 0.45) +
#   facet_grid(~ factor(group, 
#                       levels = c("triple", "other", "H3K27me3", "euchromatin")), 
#              scales = "free")
```




```{r Fig6D, fig.height= 4, fig.width= 6}
clone5_domains_damid = clone5_domains %>% left_join(analysis.mapped.integrations.df, by = "barcode")

ipr = clone5_domains_damid %>% filter(IPR == "IPR17")

PlotDataTracks(bins, 
               chr = ipr$seqname, 
               start = plyr::round_any(ipr$start, 20000)-2e6, 
               end =  plyr::round_any(ipr$start, 20000)+2e6,
               samples = c("LMNA_KO", "LBR_KO", "Clone5"),
               ipr_loc = ipr$start, cols = ko_group_colors)

```

```{r Fig6E, fig.height= 4, fig.width= 6}
clone5_domains_damid = clone5_domains %>% left_join(analysis.mapped.integrations.df, by = "barcode")

ipr = clone5_domains_damid %>% filter(IPR == "IPR2")

PlotDataTracks(bins, 
               chr = ipr$seqname, 
               start = plyr::round_any(ipr$start, 20000)-2e6, 
               end = plyr::round_any(ipr$start, 20000)+2e6,
               samples = c("LMNA_KO", "LBR_KO", "Clone5"),
               ipr_loc = ipr$start, cols = ko_group_colors)

```

```{r Fig6FG, fig.height= 6, fig.width= 6}
trip_tib_KO_domains2 <- trip_tib_KO_domains %>% 
  dplyr::select(barcode, ID, drug, KO, KO_clone, IPR, MMEJ_MMEJNHEJ, efficiency) %>%  # select columns of interest
  mutate(KO_clone = ifelse(KO == "Clone5", "wt", KO_clone)) %>% # rename so we can merge later
  group_by(barcode, ID, KO_clone, IPR, KO) %>% # grouping for means
  dplyr::summarise(MMEJ_MMEJNHEJ = mean(MMEJ_MMEJNHEJ), efficiency = mean(efficiency)) # calculate means for Clone5

mean_dam = left_join(trip_tib_KO_domains2, dam_dt[dam_dt$binsize == 20000, ], by=c('ID' = 'name', 'KO_clone' = 'KO')) # join damid table, only 20kb bins

ipr_dam = mean_dam %>% 
  filter(IPR %in% c("IPR17", "IPR2")) %>% ungroup() %>% # select IPRs of interest 
  mutate(KO_clone = ifelse(KO == "Clone5", "Clone5", KO_clone)) # rename wt to Clone5 for consistency

ggplot(ipr_dam, aes(x=MMEJ_MMEJNHEJ, y=z_score, color = KO, shape = KO)) +
  geom_point() +
  ggtitle('MMEJ:NHEJ vs DamID 20kb') +
  theme_bw() +
  theme(aspect.ratio=1) + 
  xlim(0, 0.45) +
  ylim(-2, 2) +
  facet_grid(IPR ~ .) +
  scale_color_manual(values = ko_group_colors)

# 
# ggplot(mean_dam, aes(x=efficiency, y=z_score, color=KO)) +
#     geom_point() +
#     ggtitle('efficiency vs DamID 20kb') +
#     facet_wrap(~IPR) +
#     theme_bw() +
#     theme(aspect.ratio=1)
```

## Supplementary
### Panel S6A
```{r FigS6A, fig.height=5, fig.width=7}
indel.dt = trip_tib %>% filter(drug %in% c("-", "BIX01294")) %>%
  dplyr::select(barcode, drug, MMEJ_MMEJNHEJ, domains) %>%
  spread(drug, MMEJ_MMEJNHEJ) %>%
  filter(!is.na(BIX01294),!is.na(`-`)) %>%
  mutate(BIX_ratio = BIX01294/`-`) %>%
  distinct(barcode, .keep_all = TRUE) %>%
  data.table()

dim(indel.dt)
# indel.dt = unique(tib.dt[exp_pool == "indel_mean_LBR2_ssODN_DMSO_64h",  
#                          c('exp_pool', 'barcode', 'SSTR', 'SSTR_MMEJSSTR',
#                            domains), with=F])


indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)


# Ratio of SSTR vs all indels
plot_comb_grid_beeswarm(indel.dt, opt_df,
                        domain_levels, 'log2(BIX_ratio)', min_count = 1)

k9me2 = indel.dt %>% filter(late_replicating == 0, 
                            LAD == 0,
                            H3K9me2 == 1,
                            H3K27me3 == 0) %>% 
  pull(BIX_ratio)

triple = indel.dt %>% filter(late_replicating == 1, 
                             LAD == 1,
                             H3K9me2 == 1,
                             H3K27me3 == 0) %>% 
  pull(BIX_ratio)

idomain = indel.dt %>% filter(late_replicating == 0, 
                              LAD == 0,
                              H3K9me2 == 0,
                              H3K27me3 == 0) %>% 
  pull(BIX_ratio)

wilcox.test(idomain, triple)
wilcox.test(idomain, k9me2)
```

### Panel S6B-E
BIX & GSK western blot representative figure & quantification

### Panel S6F
LMNA KO TIDE plot

### Panel S6G
LBR KO TIDE plot

### Panel S6H
LMNA KO Western blots

### Panel S6I
LBR KO Western blots

### Panel S7A
LBR
```{r FigS7A, fig.height 6, fig.width=7}
ipr = clone5_domains_damid %>% filter(IPR == "IPR17")

ko_colors = c("LMNA_KO1" = "#f9b016",
              "LMNA_KO2" = "#f9b016",
              "LMNA_KO3" = "#f9b016",
              "LMNA_KO4" = "#f9b016",
              "LBR_KO1" = "#006c68", 
              "LBR_KO2" = "#006c68",
              "LBR_KO3" = "#006c68",
              "LBR_KO4" = "#006c68",
              "Clone5" = "black")

PlotDataTracks(bins, 
               chr = ipr$seqname, 
               start = plyr::round_any(ipr$start, 20000)-2e6, 
               end = plyr::round_any(ipr$start, 20000)+2e6,
               samples = c("LMNA_KO1", "LMNA_KO2", "LMNA_KO3","LMNA_KO4",
                           "LBR_KO1", "LBR_KO2", "LBR_KO3", "LBR_KO4",
                           "Clone5"),
               ipr_loc = ipr$start, cols = ko_colors)
```

### Panel S7B
```{r FigS7B, fig.height= 6, fig.width= 7}
ipr = clone5_domains_damid %>% filter(IPR == "IPR2")

PlotDataTracks(bins, 
               chr = ipr$seqname,                
               start = plyr::round_any(ipr$start, 20000)-2e6, 
               end = plyr::round_any(ipr$start, 20000)+2e6,
               samples = c("LMNA_KO1", "LMNA_KO2", "LMNA_KO3","LMNA_KO4",
                           "LBR_KO1", "LBR_KO2", "LBR_KO3", "LBR_KO4",
                           "Clone5"),
               ipr_loc = ipr$start, cols = ko_colors)
```

### Panel S7C
```{r FigS7C, fig.height= 8, fig.width= 8}
ggplot(mean_dam, aes(x=MMEJ_MMEJNHEJ, y=z_score, color=KO, shape = KO)) +
  geom_point() +
  ggtitle('MMEJ:NHEJ vs DamID 20kb') +
  facet_wrap(~IPR) +
  theme_bw() +
  theme(aspect.ratio=1) +
  xlim(0, 0.45) +
  scale_color_manual(values = ko_group_colors)
```

# Session Info
```{r session info}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
# rm(list = ls(all.names = TRUE)) #will clear all objects includes hidden objects.
# gc() #free up memrory and report the memory usage.
```
