---
title: "Figure 4 and supplemental 4"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r}
StartTime <-Sys.time()
library(knitr)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path("figures/"))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# 6-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),3,8) 

# libraries:
library(ggplot2)
library(ggpubr)
# library(reshape2)
# library(tibble)
# library(GenomicRanges)
# library(rtracklayer)
library(corrr)
library(Hmisc)
library(ggbeeswarm)
library(RColorBrewer)
library(data.table)
library(dplyr)
# library(plyr)
library(tidyr)
# library(stringr)
# library(plotly)
library(ggpmisc)


color_redblue <- rev(brewer.pal(11,"RdBu"))
duocolor <- c("#EE756E", "#2BB6BE")
colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")
colores <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")
colori  <- c("other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "ssODN" = "#007A4C")
colora <- c("0" = "#808184", "1" = "#e1251b", "-7" = "#26419a", "-14" = "#26419a")
coloru <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "MMEJplusNHEJplusHDR" = "#EE8A2A", "HDR" = "#007A4C", "MMEJplusHDR" = "#EE8A2A")
colory <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "SSTR" = "#007A4C")
```

### Custom functions
Functions used thoughout this script.
```{r, message=FALSE, warnings=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), "_", filename)
  filename
}

CorrelationDFMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  hist.corr.list <- data[data$exp_pool==condition_name, ] %>% select(feature_name, targets) %>% as.matrix() %>% rcorr(type = "pearson")
  
  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]), 
                                corr = hist.corr.list$r[1, -1], 
                                pvalue = hist.corr.list$P[1, -1])
  
  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% filter(exp_pool==condition_name) %>% select(i, as.character(targets)) %>% as.matrix() %>% rcorr(type = "pearson")
  
    df <- data.frame(rowname = names(corr.list$r[1, -1]), 
                                corr = corr.list$r[1, -1], 
                                pvalue = corr.list$P[1, -1],
                                feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>% 
    transform(rowname = factor(rowname, levels = levels(targets)), 
              corr = corr, 
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = feature, alpha = pvalue < p.value)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab(paste("Pearson correlation with", feature_name)) +
    xlab("Chromatin Feature") + 
    ggtitle(paste("Repair pathways correlation with chromatin features in", condition_name, "exps")) +
    theme_bw(base_size = 16) +
    scale_fill_manual("legend", values = colores) +
    scale_y_continuous(breaks=seq(-1,1, .1)) + 
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}
```

# Data import
```{r, message=FALSE, warnings=FALSE}
load("/DATA/projects/DSBrepair/data/R/RSTP2_Indel_Chromatin_2kb.RData")
load("/DATA/projects/DSBrepair/data/R/RSTP2_Indel_Chromatin_2kb_all.RData")
load("/DATA/projects/DSBrepair/data/R/RSTP2_IndelRatios_Chromatin_2kb.RData")
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>% 
  filter(exp_pool == "indel_mean_LBR2_0_64h")

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool == "indel_mean_LBR2_0_64h")

mean_trip_tib_2000 <- trip_tib_2000 %>% 
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>% 
  distinct(barcode, exp, pathway, .keep_all = TRUE)

trip_tib$spin_state <- factor(trip_tib$spin_state, levels=c("Speckle",
                                                                      "Interior_Act1",
                                                                      "Interior_Act2",
                                                                      "Interior_Act3",
                                                                      "Interior_Repr1",
                                                                      "Interior_Repr2",
                                                                      "Near_Lm1",
                                                                      "Near_Lm2",
                                                                      "Near_Lm3",
                                                                      "Lamina",
                                                                      NA ))

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```

## Figures
### Panel A
```{r ratio and pathway distribution, fig.height=5, fig.width=6}
setwd("/DATA/projects/DSBrepair/scratch/")
ggplot(trip_tib, aes(log2(NHEJ_MMEJ), alpha = 0.8)) + 
  geom_density() + 
  theme_bw(base_size = 16) +
   scale_x_continuous(breaks=seq(0,5, 1)) 

# trip_tib %>% gather(other_indels, NHEJ, MMEJ, 
#                     key = "pathway", 
#                     value = 'ratios') %>% 
#   ggplot(., aes(ratios, fill = pathway, alpha = 0.8)) + 
#   geom_density() + 
#   theme_bw(base_size = 16) +
#   scale_fill_manual("legend", values = colores)
```

### Panel B
```{r pathways vs chrom, fig.height=8, fig.width=9}
chrom.mods.cols <- seq(grep("binsize", names(trip_tib))+1, grep("LAD", names(trip_tib))-1)

NHEJ.hist.corr <- CorrelationDFMean(trip_tib, "indel_mean_LBR2_0_64h", "NHEJ", targets = chrom.mods.cols) %>% 
  arrange(corr) %>% pull(rowname) %>% factor(.,levels = .)

lims = 1
colores <- c("other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a")

CorrelationPlotMean(trip_tib, "indel_mean_LBR2_0_64h", c("NHEJ", "MMEJ", "other_indels") , targets = NHEJ.hist.corr)
```

### Panel C
```{r scatterplot pathways vs lmnb1 and K4me1, fig.height=8, fig.width=12}
trip_tib %>% 
  select(barcode, MMEJ, NHEJ, other_indels, LMNB1, H3K4me1) %>% 
  group_by(barcode) %>% 
  gather(MMEJ, NHEJ, other_indels, key = "pathway", value = "ratio") %>%
  gather(LMNB1, H3K4me1, key = "feature", value = "value") %>%
  ggplot(., aes(ratio, value))  +  
  geom_point() +
  geom_smooth(method= "lm") + 
  theme_bw(base_size = 16) + 
  stat_cor(method = "pearson", label.x = .25, label.y = .7) +
  facet_wrap(feature ~ pathway, scales = "free", ncol = 3)
```

### Panel D
```{r}
het.domains <- colnames(trip_tib)[grep("K9me3_|K9me2_|K27me3_|LAD", colnames(trip_tib))]

trip_tib_domains <- trip_tib %>% select(barcode, het.domains, plasmid, MMEJ_MMEJNHEJ) %>% filter(!is.na(LAD), plasmid == "LBR2") %>% gather(het.domains, key = "dom_type", value = "domain") 

trip_tib_domains$dom_type <- factor(trip_tib_domains$dom_type, levels = c(het.domains[grep("K27me3", het.domains)], 
                                        het.domains[grep("K9me2", het.domains)], 
                                        het.domains[grep("K9me3", het.domains)], 
                                        "LAD"))

trip_tib_domains <- trip_tib_domains %>% mutate(dom_type = ifelse(grepl("K9me3", dom_type), "K9me3", 
                                                                  ifelse(grepl("K9me2", dom_type), "K9me2", 
                                                                         ifelse(grepl("K27me3", dom_type), "K27me3", 
                                                                                "LAD"))),
                                                domain = ifelse(grepl("iDomain|iAD", domain), "iDomain", 
                                                                "Domain"))

levels(trip_tib_domains$dom_type)

ggplot(trip_tib_domains, aes(dom_type, MMEJ_MMEJNHEJ, color = domain)) +  geom_quasirandom(size = 1, dodge.width = 1, varwidth = TRUE) + 
  theme_bw() + scale_color_manual(values = c("iDomain" = "gray", "Domain" = "black"))

trip_tib_domains %>% group_by(dom_type, domain) %>% dplyr::summarise(median_proportion = median(MMEJ_MMEJNHEJ))

trip_tib_domains %>% 
  # mutate(domain_type = paste0(domain, dom_type, sep = "_")) %>%
  group_by(dom_type) %>%  
  do(broom::tidy(wilcox.test(MMEJ_MMEJNHEJ ~ domain, data = .))) %>%
  ungroup() %>% 
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% # Calculate adjusted P.Val. 
  arrange(p.adj)

trip_tib_domains %>% 
  ungroup %>%
  group_by(dom_type, domain) %>%  
  dplyr::summarise(means = mean(MMEJ_MMEJNHEJ),
                   meadians = median(MMEJ_MMEJNHEJ)) %>%
  select(-meadians) %>%
  spread(domain, means) %>%
  mutate(fold_change = Domain/iDomain)

trip_tib_domains %>% 
  ungroup %>%
  group_by(dom_type, domain) %>%  
  dplyr::summarise(means = mean(MMEJ_MMEJNHEJ),
                   meadians = median(MMEJ_MMEJNHEJ)) %>%
  select(-meadians) %>%
  spread(domain, means) %>%
  mutate(fold_change = (1-Domain)/(1-iDomain))
```


## Supplementary figures
### Panel A
To be done still - I need the domains.
```{r MMEJ ratio per spin, fig.height=6, fig.width=10}
spin_color <- trip_tib %>% select(spin_state, spin_color) %>% distinct(spin_state, .keep_all = T) %>% filter(!is.na(spin_state))
colors <- spin_color$spin_color
names(colors) <- spin_color$spin_state
ggplot(filter(trip_tib, !is.na(spin_state)), aes(spin_state, MMEJ_MMEJNHEJ, color = spin_state)) + 
  geom_quasirandom(size = 3) + 
  scale_color_manual(values = colors) + 
  theme_bw(base_size = 16)
```

### Panel B
```{r MMEJ ratio per heterochomatin domain, fig.height=6, fig.width=10}
het.domains <- colnames(trip_tib)[grep("K9me3_|K9me2_|K27me3_|LAD", colnames(trip_tib))]

trip_tib_domains <- trip_tib %>% select(barcode, het.domains, MMEJ_MMEJNHEJ) %>% filter(!is.na(LAD)) %>% gather(het.domains, key = "dom_type", value = "domain") 

trip_tib_domains$dom_type <- factor(trip_tib_domains$dom_type, levels = c(het.domains[grep("K27me3", het.domains)], 
                                        het.domains[grep("K9me2", het.domains)], 
                                        het.domains[grep("K9me3", het.domains)], 
                                        "LAD"))

trip_tib_domains <- trip_tib_domains %>% mutate(dom_type = ifelse(grepl("K9me3", dom_type), "K9me3", 
                                                                  ifelse(grepl("K9me2", dom_type), "K9me2", 
                                                                         ifelse(grepl("K27me3", dom_type), "K27me3", 
                                                                                "LAD"))),
                                                domain = ifelse(grepl("iDomain|iAD", domain), "iDomain", 
                                                                "Domain"))


levels(trip_tib_domains$dom_type)
ggplot(trip_tib_domains, aes(dom_type, MMEJ_MMEJNHEJ, color = domain)) +  geom_quasirandom(size = 3, dodge.width = 1) + 
  theme_bw()
```

Panel C

# Same with all the indels
```{r fig.height=7, fig.width=12}

```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

