---
title: "Figure 7 and supplementary 7-8"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r}
StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

## Select outdir
out.dir = paste0("figures/rs", Date, "/")
dir.create(out.dir)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# INPUT DIRECTORY
in.dir.date = 20210105
in.dir = paste0("/DATA/projects/DSBrepair/data/R/rs", in.dir.date, "/")

# libraries:
library(ggplot2)
library(ggpubr)
library(tibble)
library(corrr)
library(Hmisc)
library(ggbeeswarm)
library(RColorBrewer)
library(data.table)
library(dplyr)
library(tidyr)
library(cowplot)
library(report)


colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colores <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colory <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "SSTR" = "#007A4C")

```

## Custom functions
Functions used thoughout this script.
```{r, message=FALSE, warnings=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), "_", filename)
  filename
}

CorrelationDFMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  hist.corr.list <- data[data$exp_pool==condition_name, ] %>% dplyr::select(all_of(feature_name), targets) %>% as.matrix() %>% rcorr(type = "pearson")

  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]),
                                corr = hist.corr.list$r[1, -1],
                                pvalue = hist.corr.list$P[1, -1])

  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% filter(exp_pool==condition_name) %>% dplyr::select(all_of(i), as.character(targets)) %>% as.matrix() %>% rcorr(type = "pearson")

    df <- data.frame(rowname = names(corr.list$r[1, -1]),
                                corr = corr.list$r[1, -1],
                                pvalue = corr.list$P[1, -1],
                                feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>%
    transform(rowname = factor(rowname, levels = levels(targets)),
              corr = corr,
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = feature, alpha = pvalue < p.value)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab("Pathway correlation with chromatin feature") +
    xlab("chromatin feature") +
    theme_bw(base_size = 16) +
    scale_fill_manual("legend", values = colores) +
    scale_y_continuous(breaks=seq(-1,1, .1)) +
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

cor_test <- function(y, X, left, right){
    cor_list = lapply(X, function(x){
        cor.test(x,y)
    })
    p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
    cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
    p_adj = p.adjust(p_vec)
    color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                   'None')
    importance = abs(cor_vec)
    target=factor(colnames(X), levels=rownames(clustering))
    feature_group = clustering[colnames(X), ]
    return(list(correlation=cor_vec,importance = importance, p=p_vec,
                p_adj=p_adj, target=target, color=color,
                feature_group=feature_group))
}

get_combinations <- function(domain_vec){
    n_domain = length(domain_vec)
    get_ivec <-function (last_x, n_domain){
    	if (last_x < n_domain){
    		i_vec = c(0, c((last_x+1):n_domain))
    	} else{
    		i_vec = c(0)
    	}
    }

    combine <- function(x_new, x_vec, n_domain){
        new_vec = c(x_vec, x_new)
        if (length(new_vec)==n_domain){
            return(new_vec)
        } else {
            i_vec = get_ivec(max(new_vec), n_domain)
            x_list = lapply(i_vec, combine, x_vec=new_vec,
                            n_domain=n_domain)
            return(do.call(rbind, x_list))
        }
    }

    comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)

    comb = do.call(rbind, comb_list)
    comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))

    name_list = apply(comb_df, 1, function(x){
    	name = paste(domain_vec[x], collapse='-')
    	if (name == ''){
    		name = 'iDomain'
    	}
    	return(name)
    })

    opt_list = lapply(domain_vec, function(d){
    	 apply(comb_df, 1, function(x){
    		opt = domain_vec[x]
    		return(d%in%opt)
    	})
    })

    opt_df = do.call(cbind, opt_list)
    colnames(opt_df) = domain_vec
    rownames(opt_df) = name_list

    return(opt_df)
}

test_wilcox <- function(dt, group_by='condition', y_name='MMEJ_MMEJNHEJ'){
    bc_vec = dt[,unique(barcode)]
    name1 = unlist(unique(dt[,group_by, with=F])[1])
    dt_c = dt[get(group_by)==eval(name1), ]
    dt_s = dt[get(group_by)!=eval(name1), ]
    setkey(dt_c, 'barcode')
    setkey(dt_s, 'barcode')

    w = wilcox.test(unlist(dt_c[bc_vec,..y_name]),
                    unlist(dt_s[bc_vec,..y_name]),
                    paired=T)
    return(list(p_value = w$p.value))
}



plot_comb_grid <- function(bix.dt, opt_df, domains, y_name, min_count=0,
                           max_count=Inf, group_by=NULL, color=NULL,
                           y_plabel=c(-3, 3)){

    opt_melt = reshape2::melt(opt_df)
    colnames(opt_melt) = c('group', 'domain', 'is_in')

    opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
    opt_melt$domain = factor(opt_melt$domain, levels=domains)


    group_list = lapply(rownames(opt_df), function(opt_name){
    	opt = opt_df[opt_name, ]
    	in_vec = domains[opt]
    	out_vec = domains[!opt]

    	in_sum = rowSums(bix.dt[,..in_vec])
    	out_sum = rowSums(bix.dt[,..out_vec])
    	if (length(in_vec)==0){
    		in_group = which(out_sum == 0)
    	} else if (length(out_vec) == 0){
    		in_group = which(in_sum==length(in_vec))
    	} else  {
    		in_group = which(in_sum == length(in_vec) & out_sum==0)
    	}
    	return(in_group)
    })

    names(group_list) = rownames(opt_df)
    for (opt_name in names(group_list)){
    	i_vec = group_list[[opt_name]]
    	bix.dt[i_vec, group := opt_name]
    }
    if (is.null(group_by)){
        group_count_df = data.frame(bix.dt[,table(group)])
    } else {
        group_count_df = data.frame(bix.dt[,table(group)/length(unique(get(group_by)))])
    }
    group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
    # print(group_count_df)
    group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
    group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
    group_levels =  group_count_sel$group
    opt_sel = opt_melt[opt_melt$group %in% group_levels, ]

    opt_sel$group = factor(opt_sel$group, levels=group_levels)
    # print(opt_sel)

    indel_selection = bix.dt[group%in%group_levels, ]
    indel_selection[,group:=factor(group, levels=group_levels)]


    group_count_sel$group = factor(group_count_sel$group, levels=group_levels)

    count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
                            geom_label() +
                            theme_void()



    dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    	geom_point(size=2) +
    	geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    	theme_bw() +
    	scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    	guides(color=FALSE)  +
    	geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
    			   lwd=0.5, colour="grey90") +
    	theme(axis.title=element_blank(),
    		  axis.text=element_blank(),
    		  axis.ticks=element_blank(),
    		  panel.grid.major = element_blank())

    symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                  symbols = c("****", "***", "**", "*", " "))


    if (is.null(group_by)){
      level_vec = indel_selection[, levels(group)]
      wilcox_list = lapply(level_vec[level_vec!='iDomain'],
                           function(g){
                               x = indel_selection[group==g, y_name, with=F]
                               y = indel_selection[group=='iDomain', y_name,
                                                 with=F]
                               w = wilcox.test(unlist(x),unlist(y))
                               return(data.table('group1'='iDomain', 'group2'=g,
                                                 'p'=w$p.value))
                           })
      wilcox_dt = do.call(rbind, wilcox_list)
      wilcox_dt[,p_adj := p.adjust(p)]
      wilcox_dt[,p_label := sprintf('%.3g', p_adj)]
      wilcox_dt$y.position = seq(y_plabel[1], y_plabel[2],
                                 length.out=nrow(wilcox_dt))

      comparisons = lapply(level_vec[level_vec!='iDomain'],
                           function(x){return(c('iDomain', x))})
      beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
        geom_quasirandom(width=0.3) +
        stat_summary(fun=median, fun.min = median,
                     fun.max = median,
                     geom="crossbar", width = 0.5,
                     color='red') +
        stat_pvalue_manual(wilcox_dt, label = "p_label")
    } else {
        wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                    by='group']
        wilcox_dt[,p_adj:=p.adjust(p_value)]
        wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                    symbols=symnum$symbols)]
        beeswarm = ggplot(indel_selection,
                          aes_string(x='group', y=y_name, color=group_by)) +
        			   geom_quasirandom(dodge.width=1) +
                       scale_color_manual(values=color) +
                       stat_summary(aes_string(group=group_by), fun=median,
                                    fun.min = median, fun.max = median,
                                    geom="crossbar", width = 0.5,
                                    color='red',
                                    position= position_dodge(width =1)) +
                       geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                                 inherit.aes=F)
    }
    beeswarm = beeswarm +
    			  theme_bw() +
    			  geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
    			             lwd=0.5, colour="grey90") +
    			  theme(axis.title=element_blank(),
    			        axis.text.x=element_blank(),
      				    panel.grid.major.x = element_blank(),
    				    axis.ticks.x=element_blank())


    count = data.frame(domain=gsub('_domain', '', domains),
    	               count=colSums(bix.dt[,..domains], na.rm=T))

    count$domain = factor(count$domain, levels=domain_levels)


    histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    				geom_label() +
    				theme_bw() +
    				theme(plot.background = element_blank(),
        				  panel.grid.major = element_blank(),
        				  panel.grid.minor = element_blank(),
        				  panel.border = element_blank(),
    					  axis.title=element_blank(),
    	  			      axis.text.x=element_blank(),
    	  				  axis.ticks=element_blank())

    empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
                geom_point() +
                theme_void()

    count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                               align='v', axis='lr')
    return(plot_grid(empty, beeswarm, empty, count_plot,
                     histogram, dots, nrow=3,
                     rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                     axis='tlbr'))
}

```

# Data import
```{r data loading, message=FALSE, warnings=FALSE}
trip_tib_mut_2000 = readRDS(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RDS"))
trip_tib_2000 = readRDS(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RDS"))
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>%
  filter(exp_pool  %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_NU7441_64", "mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_NU7441_64", "mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

mean_trip_tib_2000 <- trip_tib_2000 %>%
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```

# Main Figure 7

## Panel 7A

```{r Fig7A pathways vs contribution ssODN, fig.height=6, fig.width=5}
trip_tib_pathways <- trip_tib %>%
  gather(NHEJ, MMEJ, other_indels, SSTR, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

exp_order <- c("mean_RSTP2_2000_LBR2_DMSO_64",
              "mean_RSTP2_2000_LBR2_NU7441_64",
              "mean_RSTP2_2000_LBR2_DMSO_ssODN_64",
              "mean_RSTP2_2000_LBR2_NU7441_ssODN_64")

trip_tib_pathways$pathway <- factor(trip_tib_pathways$pathway, levels=c("other_indels", "MMEJ", "SSTR", "NHEJ"))
trip_tib_pathways$exp_pool <- factor(trip_tib_pathways$exp_pool, levels=exp_order)
trip_tib_pathways <- trip_tib_pathways[!is.na(trip_tib_pathways$ratios), ]

pathway_contribution <- trip_tib_pathways %>% group_by(exp_pool, pathway) %>% dplyr::summarise(pathway_contr = mean(ratios))

colores <- c("other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")

p <- ggplot(pathway_contribution, aes(exp_pool, pathway_contr, fill = pathway, order = pathway)) +
  geom_bar(stat = "identity", position = "fill")  +
  theme_bw(base_size = 16)
p + scale_fill_manual("legend", values = colores) +
  scale_x_discrete(labels=c("mean_RSTP2_2000_LBR2_DMSO_64" = "DMSO",
                            "mean_RSTP2_2000_LBR2_NU7441_64" = "ssODN + DMSO",
                            "mean_RSTP2_2000_LBR2_DMSO_ssODN_64" = "NU7441",
                            "mean_RSTP2_2000_LBR2_NU7441_ssODN_64" = "ssODN + NU7441")) +
  labs(x = element_blank(),
       y = "abundance")
```

### Figure legend
(A) Average pathway contribution across all IPRs in the cell pools, in the absence or presence of a ssODN donor, and with or without NU7441 treatment (n = 2-8). Red: +1 insertion (NHEJ); blue: -7 deletion (MMEJ); green: +2 insertion due to SSTR; black: other indels.

## Panel 7B
```{r Fig7B pathway contribution per barcode with DMSO, fig.height=4, fig.width=14}
pathway_distribution <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios))

barcode_order <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>%
  dplyr::select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("other_indels", "MMEJ", "SSTR", "NHEJ"))

trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>% ggplot(.,
       aes(factor(barcode,levels=barcode_order$barcode), efficiency, width=1)) +
  geom_bar(stat = "identity") +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("TIF") +
  xlab("IPR")

ggplot(pathway_distribution,
       aes(barcode, ratios, width=.8)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  xlab("IPR") +
  scale_fill_manual("legend", values = colores) +
  ggtitle("Pathway contribution per barcode")
```

```{r SSTR vs TIL, fig.height=5, fig.width=5}
trip_tib %>%
  filter(drug == "DMSO", ssODN == "ssODN") %>%
  ggplot(., aes(efficiency, SSTR)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  labs(title = "efficiency vs SSTR in DMSO condition",
       x = "total indel frequency",
       y = "SSTR proportion")  +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "bottom")
```



### Figure legend
(B) Pathway contributions in all individual IPRs in the presence of ssODN donor but without NU7441 (bottom panel), sorted by overall indel frequency (top panel). 

## Panel 7C

```{r Fig7C SSTR proportion per heterochomatin domain, fig.height=6, fig.width=10}
tib.dt = trip_tib %>% data.table()

domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

indel.dt = unique(tib.dt[exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64",  
                         c('exp_pool', 'barcode', 'SSTR', 'SSTR_MMEJSSTR', 'NHEJ', 'MMEJ',
                           domains), with=F])

indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)


# Ratio of SSTR vs all indels
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'SSTR', min_count = 20, y_plabel=c(0.2, 0.26))


```

### Figure legend
(C-E) Proportions of total indels repaired by SSTR (C), MMEJ (D) and NHEJ (E) in the presence of the ssODN donor, split according to heterochromatin features as indicated in the bottom panel, similar to Figure 3E. Asterisks mark p-values according to the Wilcoxon test, compared to euchromatin IPRs (most left column): * p < 0.05, ** p < 0.01, *** p < 0.001, **** p < 0.0001. 

## Panel 7D

```{r Fig7D MMEJ proportion per heterochomatin domain, fig.height=6, fig.width=10}
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'MMEJ', min_count = 20, y_plabel=c(.25, .3))
```

## Panel 7E

```{r Fig7E NHEJ proportion per heterochomatin domain, fig.height=6, fig.width=10}
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'NHEJ', min_count = 20, y_plabel=c(.8, .9))
```

# Supplementary Figure 7
## Panel S7A
Schematic of the ssODN
### Figure legend
(A) Schematic of the strategy to probe SSTR simultaneously with NHEJ and MMEJ. Prior to Cas9 activation, the ssODN is co-transfected with a plasmid that encodes the LBR2 sgRNA. The ssODN (black bar) covers the reporter sequence but not the barcode, and contains a 2 bp insertion (green) at the PAM. 


## Panel S7B
```{r FigS7B indel distribution ssODN, fig.height=4, fig.width=6}
SSTR_indels = filter(trip_tib_mut, exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

SSTR_indels %>% 
  filter(drug == "DMSO") %>%
  ggplot(., aes(mutation, freq, color = color)) +
  geom_quasirandom() +
  theme_bw(base_size = 16) +
  theme(legend.position = c(0.2, 0.6)) +
  scale_color_manual("legend", values = colory) +
  facet_grid(. ~ drug)

SSTR_indels %>%
  group_by(mutation, drug) %>%
  dplyr::summarise(median = median(freq)) %>%
  arrange(dplyr::desc(median))

SSTR_indels %>%
  group_by(drug) %>%
  distinct(barcode, drug) %>% 
  dplyr::summarise(n_barcodes = n())

trip_tib_mut %>%
  distinct(exp_pool, nexp)

```

### Figure legend
(B) Indel frequencies generated by NHEJ, MMEJ and SSTR in 965 IPRs in the two cell pools, 64 hours after Cas9 activation (average of two replicate experiments). Left panel with DMSO, right panel with 1mM NU7441. N = 2-3. 

## Panel S7C
```{r FigS7C scatter NHEJ with and without ssODN, fig.height=5, fig.width=5}
trip_lbr2_ssODN_NHEJ <- trip_tib %>% filter(drug == "DMSO") %>% dplyr::select(barcode, ssODN, NHEJ, late_replicating) %>% spread(ssODN, NHEJ)

ggplot(trip_lbr2_ssODN_NHEJ, aes(`-`, ssODN)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_abline(slope = 1)  +
  geom_smooth(method = glm) +
  xlim(.3, .82) + ylim(.3, .82)  +
  labs(title = "NHEJ - vs NHEJ + ssODN in DMSO condition",
       x = "NHEJ -",
       y = "NHEJ ssODN") +
  stat_cor(method = "pearson")
```

### Figure legend
(C-D) Comparison of +1 (NHEJ; panel C) and -7 (MMEJ, panel D) indel frequencies in all IPRs in cell pools in the presence (+) or absence (-) of the ssODN. Black line: diagonal. 


## Panel S7D
```{r FigS7D scatters MMEJ ssODN vs MMEJ - DMSO, fig.height=5, fig.width=5}
formula <- y ~ x

trip_lbr2_ssODN_MMEJ <- trip_tib %>% filter(drug == "DMSO") %>% dplyr::select(barcode, ssODN, MMEJ, late_replicating) %>% spread(ssODN, MMEJ)

ggplot(trip_lbr2_ssODN_MMEJ, aes(`-`, ssODN)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_abline(slope = 1)  +
  geom_smooth(method = glm) +
  xlim(0,.45) + ylim(0,.45)  +
  labs(title = "MMEJ - vs MMEJ + ssODN in DMSO condition",
       x = "MMEJ -",
       y = "MMEJ ssODN") +
  stat_cor(method = "pearson", label.x = 0, label.y = .3)
```

###  Figure legend

## Panel S7E

```{r FigS7E MMEJ vs SSTR with DMSO, fig.height=5, fig.width=5}
trip_tib %>%
  filter(drug == "DMSO", ssODN == "ssODN") %>%
  ggplot(., aes(MMEJ, SSTR)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  xlim(0,.3) + ylim(0,.3)  +
  labs(title = "MMEJ vs SSTR in DMSO condition",
       x = "MMEJ",
       y = "SSTR") +
  stat_cor(method = "pearson",label.x.npc = "left", label.y.npc = "top")
```
# Bibliography
```{r citations}
cite_packages()
```


# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```
