---
title: "Figure 4 and supplemental 4"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r}
StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8) 

## Select outdir
out.dir = paste0("figures/rs", Date, "/")
dir.create(out.dir)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# INPUT DIRECTORY
in.dir.date = 20191230
in.dir = paste0("/DATA/projects/DSBrepair/data/R/rs", in.dir.date, "/")

# libraries:
library(ggplot2)
library(ggpubr)
# library(reshape2)
# library(tibble)
# library(GenomicRanges)
# library(rtracklayer)
library(corrr)
library(Hmisc)
library(ggbeeswarm)
library(RColorBrewer)
library(data.table)
library(dplyr)
# library(plyr)
library(tidyr)
# library(stringr)
# library(plotly)
library(ggpmisc)

library(stats)
library(glmnet)


library(cowplot)

library(mda)
library(earth)
library(yaml)
library(vip)
library(caret)
library(scales)
library(glmnet)
library(gridExtra)
library(ggcorrplot)
library(bnlearn)
library(pheatmap)
library(ppcor)
library(parallel)


color_redblue <- rev(brewer.pal(11,"RdBu"))
duocolor <- c("#EE756E", "#2BB6BE")
colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")
colores <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")
colori  <- c("other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "ssODN" = "#007A4C")
colora <- c("0" = "#808184", "1" = "#e1251b", "-7" = "#26419a", "-14" = "#26419a")
coloru <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "MMEJplusNHEJplusHDR" = "#EE8A2A", "HDR" = "#007A4C", "MMEJplusHDR" = "#EE8A2A")
colory <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "SSTR" = "#007A4C")

measure_color = c(NHEJ='#e1251b', MMEJ='#26419a', SSTR='#007A4C', None='grey')

method_color = c(MARS='#00bfc4', LASSO='#f8766d')
lad_color <- c(LAD="#802cdc", iLAD="#dfaf2f")
sign_color <- c(down="#AA6375", up="#7B9A36")

feature_fill = c(insulator="#a6b5a3", repressive="#304441", euchromatin="#cda4cc",
                 transcribing="#ED9114", HDAC="#aac9e0", accessibility='#cebc85',
                 methylation='#7dc98f')

feature_color = c(insulator="black", repressive="white", euchromatin="black",
                  transcribing="black", HDAC="black", accessibility='black',
                  methylation='black')
```

### Custom functions
Functions used thoughout this script.
```{r, message=FALSE, warnings=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), "_", filename)
  filename
}

CorrelationDFMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  hist.corr.list <- data[data$exp_pool==condition_name, ] %>% dplyr::select(feature_name, targets) %>% as.matrix() %>% rcorr(type = "pearson")
  
  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]), 
                                corr = hist.corr.list$r[1, -1], 
                                pvalue = hist.corr.list$P[1, -1])
  
  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% filter(exp_pool==condition_name) %>% dplyr::select(i, as.character(targets)) %>% as.matrix() %>% rcorr(type = "pearson")
  
    df <- data.frame(rowname = names(corr.list$r[1, -1]), 
                                corr = corr.list$r[1, -1], 
                                pvalue = corr.list$P[1, -1],
                                feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>% 
    transform(rowname = factor(rowname, levels = levels(targets)), 
              corr = corr, 
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = feature, alpha = pvalue < p.value)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab(paste("Pearson correlation with", feature_name)) +
    xlab("Chromatin Feature") + 
    ggtitle(paste("Repair pathways correlation with chromatin features in", condition_name, "exps")) +
    theme_bw(base_size = 16) +
    scale_fill_manual("legend", values = colores) +
    scale_y_continuous(breaks=seq(-1,1, .1)) + 
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

bayesian_boot_func <- function(Xy, y_name, blacklist){
	n = nrow(Xy)
	boot.idx <- sample(1 : n, size = n, replace = TRUE)
	Xy.boot <- Xy[boot.idx, ]
	hc.boot = hc(Xy.boot, blacklist=blacklist)
	arc_matrix = arcs(hc.boot)
	selection = arc_matrix[arc_matrix[,2]==y_name,1]
	pcor_list = lapply(selection, function(sel){
		other = selection[selection!=sel]
		pcor.test(Xy.boot[,..sel], Xy.boot[,..y_name],
			      Xy.boot[,..other])$estimate
	})

	result_vec = rep(0,ncol(Xy)-1)
	names(result_vec) = colnames(Xy)[colnames(Xy)!=y_name]
	result_vec[selection] = unlist(pcor_list)
	return(result_vec)
}


bootstrap_bayes <- function(Xy, y_name, blacklist, nboot=1000, ncores=20){
	baye_boot_list = mclapply(1:nboot, function(i){
		bayesian_boot_func(Xy, y_name, blacklist)
	}, mc.cores=ncores)
	baye_boot.df = do.call(cbind, baye_boot_list)
	sel_baye = rowSums(baye_boot.df!=0) / nboot

	z_vec = apply(baye_boot.df,1, function(x){mean(x)/sd(x)})

	med_vec = apply(baye_boot.df, 1, function(x){
		median(x[x!=0])
	})
	pos_ratio_vec = apply(baye_boot.df, 1, function(x){
		length(which(x>0))/length(which(x!=0))
	})

	baye.dt = data.table(var = names(sel_baye), importance=sel_baye, z=z_vec,
                         median=med_vec, pos_ratio=pos_ratio_vec)

	return(baye.dt)
}



bootstrap_pcorsign_func <- function(X, y, all){
	n = nrow(X)
	boot.idx <- sample(1 : n, size = n, replace = TRUE)
	X.boot <- X[boot.idx, ]
	y.boot <- y[boot.idx]
	fit <- earth(x = X.boot, y = y.boot)
	nterms = length(fit$selected.terms)

	mm = data.frame(y = y.boot,
		            model.matrix(fit, X.boot, which.terms=1:nrow(fit$dirs))[,-1])

	coef_matrix = apply(fit$prune.terms[2:nterms, ], 1, function(x){
		xnn = x[x>0]
		dirs = fit$dirs[xnn, ,drop=F]
        count = colSums(dirs>0)
        selection = names(count)[count>0]
        if (length(selection) > 1){
            pcor_list = lapply(selection, function(sel){
        		other = selection[selection!=sel]
        		pcor.test(X.boot[,..sel], y.boot,
        			      X.boot[,..other])$estimate
        	})
        } else if (length(selection) == 1){
            pcor_list = list(cor(unlist(X.boot[,..selection]), y.boot))
        }
		result = rep(0, ncol(X))
		names(result) = colnames(X)
        if (length(selection) > 0){
            result[selection] = unlist(pcor_list)
        }
		return(result)
	})
	if (all){
		imp = evimp(fit)[,'nsubsets']
		names(imp) = gsub('-unused', '', names(imp))
		imp_vec = rep(0,ncol(X))
		names(imp_vec) = colnames(X)
		imp_vec[names(imp)] = imp / nterms
	} else {
		imp_vec = rowSums(coef_matrix!=0) / nterms
	}

	# result[i_vec] = result[i_vec]/2
	return(list(importance=imp_vec, coef=coef_matrix))
}



bootstrap_mars_pcor <-function(X, y, nboot=1000, all=T, ncores=20){
	mars_boot_list = mclapply(1:nboot, function(i){
		b_lasso = bootstrap_pcorsign_func(X, y, all)
	}, mc.cores=ncores)

	mars_boot.df = do.call(cbind, lapply(mars_boot_list,
		                                 function(x){x[['importance']]}))
	sel_mars = rowSums(mars_boot.df) / nboot

	coef_matrix = do.call(cbind, lapply(mars_boot_list,
		                                function(x){x[['coef']]}))

	z_vec = apply(coef_matrix, 1, function(x){
		mean(x) / sd(x)
	})
	med_vec = apply(coef_matrix, 1, function(x){
		median(x[x!=0])
	})
	pos_ratio_vec = apply(coef_matrix, 1, function(x){
		length(which(x>0))/length(which(x!=0))
	})
	mars.dt = data.table(var = names(sel_mars), importance=sel_mars,
	                     z=z_vec[names(sel_mars)], median=med_vec,
						 pos_ratio=pos_ratio_vec)

	return(mars.dt)
}




plot_MARS_bayes <- function(left_set, right_set, none_label='ambiguous',
                            top=NULL, left_title='Bayesian Network',
                            right_title='MARS'){

	color_labels = left_set[,levels(color)]
	color_labels[color_labels == 'None'] = none_label

	g_legend<-function(a.gplot){
	  tmp <- ggplot_gtable(ggplot_build(a.gplot))
	  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
	  legend <- tmp$grobs[[leg]]
	  return(legend)
	}
	break_vec = seq(0,1,0.2)
	minor_break_vec = seq(0.1,0.9,0.2)

	g_left <- ggplot(data = left_set, aes(x = target, y = importance, fill=color)) +
	  geom_bar(stat = "identity") +
	  theme_bw() +
	  ggtitle(left_title) +
	  theme(axis.title.x = element_blank(),
	        axis.title.y = element_blank(),
	        axis.text.y = element_blank(),
	        axis.ticks.y = element_blank(),
	        plot.margin = unit(c(1,0,1,0), "mm"),
		    plot.title = element_text(hjust = 1)) +
	  scale_fill_manual(values=measure_color,
	                    labels=color_labels) +
	  scale_y_reverse(limits=c(1.1,0), expand=c(0,0),
	                  breaks=break_vec,
				      minor_breaks=minor_break_vec) +
	  coord_flip()


	g_right <- ggplot(right_set, aes(x = target, y = importance, fill=color)) +
	  geom_bar(stat = "identity") +
	  theme_bw() +
	  ggtitle(right_title) +
	  scale_fill_manual(values=c(measure_color, None='gray'),
	        		    labels=color_labels) +
	  theme(axis.title.x = element_blank(),
	        axis.title.y = element_blank(),
	        axis.text.y = element_blank(),
			axis.ticks.y = element_blank(),
	        plot.margin = unit(c(1,0,1,-1), "mm")) +
	  scale_y_continuous(limits=c(0,1.1),
		                 expand=c(0,0),
	                     breaks=break_vec,
				         minor_breaks=minor_break_vec) +
	  coord_flip() +
	  guides(fill=F)

	g_mid<-ggplot(left_set, aes(x=1,y=target, label=target, fill=feature_group,
		          color=feature_group, yend=target)) +
	  geom_label(show.legend = F) +
	  geom_segment(aes(x=0.94,xend=0.945), color='black')+
	  geom_segment(aes(x=1.056,xend=1.061), color='black')+
	  ggtitle("")+
	  ylab(NULL)+
	  scale_color_manual(values=feature_color) +
	  scale_fill_manual(values=feature_fill) +
	  scale_x_continuous(expand=c(0,0),limits=c(0.94,1.065)) +
	  theme(axis.title=element_blank(),
	        panel.grid=element_blank(),
	        axis.text.y=element_blank(),
	        axis.ticks.y=element_blank(),
	        panel.background=element_blank(),
	        axis.text.x=element_text(color=NA),
	        axis.ticks.x=element_line(color=NA),
	        plot.margin = unit(c(1,-1,1,-1), "mm"))

	gg_left <- ggplot_gtable(ggplot_build(g_left + guides(fill=F)))
	gg_right <- ggplot_gtable(ggplot_build(g_right))
	gg_mid <- ggplot_gtable(ggplot_build(g_mid + guides(fill=F, color=F)))

	measure_legend <- g_legend(g_left + guides(fill=guide_legend(title="preferred\npathway")))
	feature_legend <- g_legend(g_mid +
		                       geom_bar(stat='identity') +
							   guides(fill=guide_legend(title="feature group"),
	                                  color=F))

	layout = rbind(c(rep(1,3),2,rep(3,3),4),
	               c(rep(1,3),2,rep(3,3),5))

	return(grid.arrange(gg_left,gg_mid, gg_right, measure_legend, feature_legend,
		                layout_matrix=layout, top=top))
}




clustering = data.frame(row.names=c('LMNB1', 'late_replicating', 'H3K9me2',
                                    'H3K9me3', 'H3K27me3', 'EZH2', 'CTCF', 'SMC3',
							        'HDAC3', 'HDAC2', 'HDAC1', 'H3K4me1',
								    'H3K4me2', 'H3K4me3', 'H3K27ac',
								    'H4K5acK8ac', 'H2AFZ', 'DNAse', 'Dam', 'm5C',
								    'H3K79me2', 'TTseq', 'H3K36me3', 'POL2AS2',
								    'POL2'),
					   group=factor(c(rep('repressive', 6), rep('insulator', 2),
				                      rep('HDAC', 3), rep('euchromatin', 6),
                                      rep('accessibility', 2), 'methylation',
						              rep('transcribing', 5)),
									levels=c('transcribing', 'accessibility',
                                             'methylation', 'euchromatin',
								             'HDAC', 'insulator', 'repressive')))



cor_test <- function(y, X, left, right){
    cor_list = lapply(X, function(x){
        cor.test(x,y)
    })
    p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
    cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
    p_adj = p.adjust(p_vec)
    color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                   'None')
    importance = abs(cor_vec)
    target=factor(colnames(X), levels=rownames(clustering))
    feature_group = clustering[colnames(X), ]
    return(list(correlation=cor_vec,importance = importance, p=p_vec,
                p_adj=p_adj, target=target, color=color,
                feature_group=feature_group))
}
```

# Data import
```{r, message=FALSE, warnings=FALSE}
load(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RData"))
load(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RData"))
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>% 
  filter(exp_pool == "indel_mean_LBR2_DMSO_64h")

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool == "indel_mean_LBR2_DMSO_64h")

mean_trip_tib_2000 <- trip_tib_2000 %>% 
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>% 
  distinct(barcode, exp, pathway, .keep_all = TRUE)

trip_tib$spin_state <- factor(trip_tib$spin_state, levels=c("Speckle",
                                                                      "Interior_Act1",
                                                                      "Interior_Act2",
                                                                      "Interior_Act3",
                                                                      "Interior_Repr1",
                                                                      "Interior_Repr2",
                                                                      "Near_Lm1",
                                                                      "Near_Lm2",
                                                                      "Near_Lm3",
                                                                      "Lamina",
                                                                      NA ))

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```

## Figures

### Panel A
```{r individual barcodes, fig.height=7, fig.width=6}
set.seed(2)
barcodes <- trip_tib_mut %>% filter(nexp > 5) %>% pull(barcode) %>% unique()
barcodeset <- sample(barcodes, 5)

NHEJratio <- filter(trip_tib_2000, barcode %in% barcodeset) %>% 
  distinct(barcode, .keep_all = T) %>% 
  dplyr::select(barcode, NHEJ_MMEJ, efficiency) %>% mutate(label_efficiency = paste0("Efficiency = ", round(efficiency, 2),"%"), 
                                                    label_ratio = paste0("Ratio = ", round(NHEJ_MMEJ, 2)),
                                                    label = paste0("IPR ", seq(1, 5)))

colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")

p <- ggplot(filter(trip_tib_mut, barcode %in% barcodeset, mutation != "ssODN"), 
            aes(mutation, freq)) + 
  geom_bar(stat = "identity", aes(fill = color)) + 
  scale_fill_manual("legend", values = colore) + 
  theme_bw() +
  
  geom_text(data = NHEJratio, aes(x = 5, y = .7, label = label_efficiency)) +
  geom_text(data = NHEJratio, aes(x = 5, y = .6, label = label_ratio))
p + facet_wrap( ~ barcode, ncol = 2)
```

### Panel B
```{r pathway contribution, fig.height=6, fig.width=10}
pathway_distribution <- trip_tib %>% filter(exp_pool == "indel_mean_LBR2_0_64h") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels, key = "pathway", value = 'ratios') %>% arrange(desc(ratios))
barcode_order <- pathway_distribution %>% filter(pathway == "NHEJ")  %>% arrange(desc(ratios)) %>% select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("SSTR", "MMEJ", "other_indels",  "NHEJ"))

ggplot(pathway_distribution,
       aes(barcode, ratios, width=1)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  scale_fill_manual("legend", values = colores) +
  ggtitle("Pathway contribution per barcode")
``` 

### Panel C
```{r ratio and pathway distribution, fig.height=3, fig.width=4}
ggplot(trip_tib, aes(log2(NHEJ_MMEJ), alpha = 0.8)) + 
  geom_density(fill = "black") + 
  theme_bw(base_size = 16) +
   scale_x_continuous(breaks=seq(0,5, 1)) 

# trip_tib %>% gather(other_indels, NHEJ, MMEJ, 
#                     key = "pathway", 
#                     value = 'ratios') %>% 
#   ggplot(., aes(ratios, fill = pathway, alpha = 0.8)) + 
#   geom_density() + 
#   theme_bw(base_size = 16) +
#   scale_fill_manual("legend", values = colores)
```

### Panel D
```{r pathways vs chrom, fig.height=8, fig.width=9}
chrom.mods.cols <- seq(grep("binsize", names(trip_tib))+1, grep("H3K27me3_domain", names(trip_tib))-1)

NHEJ.hist.corr <- CorrelationDFMean(trip_tib, "indel_mean_LBR2_DMSO_64h", "NHEJ", targets = chrom.mods.cols) %>% 
  arrange(corr) %>% pull(rowname) %>% factor(.,levels = .)

lims = 1
colores <- c("other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a")

CorrelationPlotMean(trip_tib, "indel_mean_LBR2_DMSO_64h", c("NHEJ", "MMEJ", "other_indels") , targets = NHEJ.hist.corr)
```



### Panel E
```{r MMEJ ratio per heterochomatin domain, fig.height=6, fig.width=10}
het.domains <- colnames(trip_tib)[grep("K9me3_|K9me2_|K27me3_|LAD|replicating_domain", colnames(trip_tib))]

trip_tib_domains <- trip_tib %>% dplyr::select(barcode, het.domains, MMEJ_MMEJNHEJ) %>% filter(!is.na(LAD)) %>% gather(het.domains, key = "dom_type", value = "domain") 

trip_tib_domains$dom_type <- factor(trip_tib_domains$dom_type, levels = c(het.domains[grep("K27me3", het.domains)], 
                                        het.domains[grep("K9me2", het.domains)], 
                                        het.domains[grep("K9me3", het.domains)], 
                                        het.domains[grep("repli", het.domains)],
                                        "LAD"))

trip_tib_domains <- trip_tib_domains %>% mutate(dom_type = ifelse(grepl("K9me3", dom_type), "H3K9me3", 
                                                                  ifelse(grepl("K9me2", dom_type), "H3K9me2", 
                                                                         ifelse(grepl("K27me3", dom_type), "H3K27me3", 
                                                                                ifelse(grepl("replic", dom_type), "RepTime", 
                                                                                       "LAD")))),
                                                domain = ifelse(grepl("iDomain|iAD", domain), "iDomain", 
                                                                "Domain"))

trip_tib_domains$dom_type <- factor(trip_tib_domains$dom_type, levels = c("H3K9me3", "H3K9me2", "H3K27me3", "RepTime", "LAD"))
levels(trip_tib_domains$dom_type)

ggplot(trip_tib_domains, aes(dom_type, MMEJ_MMEJNHEJ, color = domain)) +  
  geom_quasirandom(size = 1, dodge.width = .6, varwidth = TRUE) + 
  theme_bw(base_size = 16) + 
  scale_color_manual(values = c("iDomain" = "gray", "Domain" = "black"))  +
  stat_summary(fun.y=median, fun.ymin = median, fun.ymax = median, geom="crossbar", width = 0.5)

trip_tib_domains %>% group_by(dom_type, domain) %>% dplyr::summarise(median_proportion = median(MMEJ_MMEJNHEJ))

trip_tib_domains %>% 
  # mutate(domain_type = paste0(domain, dom_type, sep = "_")) %>%
  group_by(dom_type) %>%  
  do(broom::tidy(wilcox.test(MMEJ_MMEJNHEJ ~ domain, data = .))) %>%
  ungroup() %>% 
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>% # Calculate adjusted P.Val. 
  arrange(p.adj)

trip_tib_domains %>% 
  ungroup %>%
  group_by(dom_type, domain) %>%  
  dplyr::summarise(means = mean(MMEJ_MMEJNHEJ),
                   meadians = median(MMEJ_MMEJNHEJ)) %>%
  dplyr::select(-means) %>%
  spread(domain, meadians) %>%
  mutate(fold_change = Domain/iDomain)

trip_tib_domains %>% 
  ungroup %>%
  group_by(dom_type, domain) %>%  
  dplyr::summarise(means = mean(MMEJ_MMEJNHEJ),
                   meadians = median(MMEJ_MMEJNHEJ)) %>%
  dplyr::select(-means) %>%
  spread(domain, meadians) %>%
  mutate(fold_change = (1-Domain)/(1-iDomain))
```


## Supplementary figures
### Panel A
```{r scatterplot pathways vs lmnb1 and K4me1, fig.height=8, fig.width=12}
trip_tib %>% 
  dplyr::select(barcode, MMEJ, NHEJ, other_indels, LMNB1, H3K4me1) %>% 
  group_by(barcode) %>% 
  gather(MMEJ, NHEJ, other_indels, key = "pathway", value = "ratio") %>%
  gather(LMNB1, H3K4me1, key = "feature", value = "value") %>%
  ggplot(., aes(ratio, value))  +  
  geom_point() +
  geom_smooth(method= "lm") + 
  theme_bw(base_size = 16) + 
  stat_cor(method = "pearson", label.x = .25, label.y = .7) +
  facet_wrap(feature ~ pathway, scales = "free", ncol = 3)
```

```{r mars bayes, fig.height=8, fig.width=8}
chip.dt = readRDS('/DATA/projects/DSBrepair/data/R/cl20191217_ChIP_mean_selection.RDS')
indel.dt = data.table(trip_tib_2000)

indel_selection_64.dt = unique(indel.dt[exp=='indel_mean_LBR2_DMSO_64h',
			                         	c('barcode', 'MMEJ_MMEJNHEJ')])

indel_chip.dt = merge(indel_selection_64.dt, chip.dt,
                      by.x='barcode', by.y='ID')
xy2000_64 = indel_chip.dt[, -c('barcode', 'binsize', 'pool')]

cor_vec = xy2000_64[,cor(MMEJ_MMEJNHEJ, .SD[,-c('MMEJ_MMEJNHEJ')])][1,]

cor_set_64 = xy2000_64[,cor_test(MMEJ_MMEJNHEJ, .SD[,-c('MMEJ_MMEJNHEJ')],
                                 'MMEJ', 'NHEJ')]


blacklist = data.frame(from='MMEJ_MMEJNHEJ',
                       to=grep('MMEJ_MMEJNHEJ', colnames(xy2000_64),invert=T, value=T))

boot = boot.strength(xy2000_64, algorithm='hc', R=200,
                     algorithm.args=list(blacklist=blacklist))

bayes_set_64 = bootstrap_bayes(xy2000_64, 'MMEJ_MMEJNHEJ', blacklist, 1000)
bayes_set_64[,color:=ifelse(pos_ratio>0.7, 'MMEJ',
                        ifelse(pos_ratio < 0.3, 'NHEJ', 'None'))]
bayes_set_64[,feature_group := clustering[var,]]
bayes_set_64[,target:=factor(var, levels=rownames(clustering))]
bayes_set_64[,color:=factor(color, levels=c('MMEJ', 'NHEJ', 'None'))]

x2000_64 = xy2000_64[, -c('MMEJ_MMEJNHEJ')]
y2000_64 = xy2000_64[, c('MMEJ_MMEJNHEJ')]


# full_set = bootstrap_mars(x2000_64, y2000_64$MMEJ_MMEJNHEJ, nboot=200)
# full_set[,sign:=sign_vec[var]]
# full_set[,color:=ifelse(sign>0, 'up', 'down')]
# full_set[,signed_importance:=importance * sign]
# full_set[,target:=factor(var, levels=var[order(importance, decreasing=F)])]
MARS_set_64 = bootstrap_mars_pcor(x2000_64, y2000_64$MMEJ_MMEJNHEJ, nboot=1000)
MARS_set_64[,color:=ifelse(pos_ratio>0.7, 'MMEJ',
                        ifelse(pos_ratio < 0.3, 'NHEJ', 'None'))]
MARS_set_64[,feature_group := clustering[var,]]
MARS_set_64[,target:=factor(var, levels=rownames(clustering))]
MARS_set_64[,color:=factor(color, levels=c('MMEJ', 'NHEJ', 'None'))]



bayes_set_64[,method:='bayes']
MARS_set_64[,method:='MARS']

full_set_64 = rbind(bayes_set_64, MARS_set_64)

full_set_64[,signed_importance:=ifelse(method=='bayes', -importance, importance)]

plot_MARS_bayes(bayes_set_64, MARS_set_64, none_label = 'ambiguous',
	            top='MMEJ / (MMEJ + NHEJ) 64h feature importance')
```


To be done still - I need the domains.
```{r MMEJ ratio per spin, fig.height=6, fig.width=10}
spin_color <- trip_tib %>% dplyr::select(spin_state, spin_color) %>% distinct(spin_state, .keep_all = T) %>% filter(!is.na(spin_state))
colors <- spin_color$spin_color
names(colors) <- spin_color$spin_state
ggplot(filter(trip_tib, !is.na(spin_state)), aes(spin_state, MMEJ_MMEJNHEJ, color = spin_state)) + 
  geom_quasirandom(size = 3) + 
  scale_color_manual(values = colors) + 
  theme_bw(base_size = 16)  +
  stat_summary(fun.y=median, fun.ymin = median, fun.ymax = median, geom="crossbar", width = 0.5)
```

# Same with all the indels
```{r fig.height=7, fig.width=12}

```

# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```

