---
title: "Figure 4 and supplementary figure 4"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r knitr opt}
StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

## Select outdir
out.dir = paste0("figures/rs", Date, "/")
dir.create(out.dir)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# INPUT DIRECTORY
in.dir.date = 20201213
in.dir = paste0("/DATA/projects/DSBrepair/data/R/rs", in.dir.date, "/")

# libraries:
library(ggplot2)
library(ggpubr)
library(Hmisc)
library(ggbeeswarm)
library(data.table)
library(dplyr)
library(tidyr)
library(broom)
library(cowplot)
library(report)
library(purrr)
library(viridis)

## Colors used in this script
colores <- c("MMEJ_MMEJNHEJ" = "black")
colore <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#e1251b", "MMEJ" = "#26419a", "SSTR" = "#007A4C")
chrom_colors = c( c("euchromatin" = "#F7941D", "other" = "#838687",  "H3K27me3" = "#D21F8A", "triple" = "#662D91"))
```

## Custom functions
Functions used thoughout this script.
```{r functions, message=FALSE, warnings=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), "_", filename)
  filename
}

CorrelationDFMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  hist.corr.list <- data[data$exp_pool==condition_name, ] %>% dplyr::select(feature_name, targets) %>% as.matrix() %>% rcorr(type = "spearman")
  
  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]),
                                corr = hist.corr.list$r[1, -1],
                                pvalue = hist.corr.list$P[1, -1])
  
  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, condition_name, feature_name, targets, p.value=0.001, alfa = c(0.5, 1)){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% filter(exp_pool==condition_name) %>% dplyr::select(i, as.character(targets)) %>% as.matrix() %>% rcorr(type = "spearman")
    
    df <- data.frame(rowname = names(corr.list$r[1, -1]),
                     corr = corr.list$r[1, -1],
                     pvalue = corr.list$P[1, -1],
                     feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>%
    transform(rowname = factor(rowname, levels = levels(targets)),
              corr = corr,
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = feature)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab(paste("Spearman correlation with MMEJ:NHEJ balance")) +
    xlab("Chromatin Feature") +
    # ggtitle(paste("Repair pathways correlation with chromatin features in", condition_name, "exps")) +
    theme_bw(base_size = 16) +
    scale_fill_manual("legend", values = colores) +
    scale_y_continuous(breaks=seq(-1,1, .1)) +
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

clustering = data.frame(row.names=c('LMNB1', 'late_replicating', 'H3K9me2',
                                    'H3K27me3', 'EZH2', 'CTCF', 'SMC3',
                                    'HDAC3', 'HDAC2', 'HDAC1', 'H3K4me1',
                                    'H3K4me2', 'H3K4me3', 'H3K27ac',
                                    'H4K5acK8ac', 'H2AFZ', 'DNAse', 'Dam', 
                                    'm5C', 'H3K79me2', 'TTseq', 'H3K36me3', 
                                    'POL2AS2', 'POL2'),
                        group=factor(c(rep('repressive', 5), rep('insulator', 2),
                                       rep('HDAC', 3), rep('euchromatin', 6),
                                       rep('accessibility', 2), 'methylation',
                                       rep('transcribing', 5)),
                                     levels=c('transcribing', 'accessibility',
                                              'methylation', 'euchromatin',
                                              'HDAC', 'insulator', 'repressive')))

cor_test <- function(y, X, left, right){
  cor_list = lapply(X, function(x){
    cor.test(x,y)
  })
  p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
  cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
  p_adj = p.adjust(p_vec)
  color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                 'None')
  importance = abs(cor_vec)
  target=factor(colnames(X), levels=rownames(clustering))
  feature_group = clustering[colnames(X), ]
  return(list(correlation=cor_vec,importance = importance, p=p_vec,
              p_adj=p_adj, target=target, color=color,
              feature_group=feature_group))
}

get_combinations <- function(domain_vec){
  n_domain = length(domain_vec)
  get_ivec <-function (last_x, n_domain){
    if (last_x < n_domain){
      i_vec = c(0, c((last_x+1):n_domain))
    } else{
      i_vec = c(0)
    }
  }
  
  combine <- function(x_new, x_vec, n_domain){
    new_vec = c(x_vec, x_new)
    if (length(new_vec)==n_domain){
      return(new_vec)
    } else {
      i_vec = get_ivec(max(new_vec), n_domain)
      x_list = lapply(i_vec, combine, x_vec=new_vec,
                      n_domain=n_domain)
      return(do.call(rbind, x_list))
    }
  }
  
  comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)
  
  comb = do.call(rbind, comb_list)
  comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))
  
  name_list = apply(comb_df, 1, function(x){
    name = paste(domain_vec[x], collapse='-')
    if (name == ''){
      name = 'iDomain'
    }
    return(name)
  })
  
  opt_list = lapply(domain_vec, function(d){
    apply(comb_df, 1, function(x){
      opt = domain_vec[x]
      return(d%in%opt)
    })
  })
  
  opt_df = do.call(cbind, opt_list)
  colnames(opt_df) = domain_vec
  rownames(opt_df) = name_list
  
  return(opt_df)
}

plot_comb_grid <- function(indel.dt, opt_df, domains, y_name, min_count=0,
                           max_count=Inf, group_by=NULL, color=NULL, y_plabel=c(0.15, 0.25)){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(indel.dt[,..in_vec])
    out_sum = rowSums(indel.dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    indel.dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(indel.dt[,table(group)])
  } else {
    group_count_df = data.frame(indel.dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = indel.dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]
  
  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    theme_bw() +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color=FALSE)  +
    geom_vline(xintercept=seq(1.5, length(unique(indel.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", ""))
  
  if (is.null(group_by)){
    level_vec = indel_selection[, levels(group)]
    wilcox_list = lapply(level_vec[level_vec!='iDomain'],
                         function(g){
                           x = indel_selection[group==g, y_name, with=F]
                           y = indel_selection[group=='iDomain', y_name,
                                               with=F]
                           w = wilcox.test(unlist(x),unlist(y))
                           return(data.table('group1'='iDomain', 'group2'=g,
                                             'p'=w$p.value))
                         })
    wilcox_dt = do.call(rbind, wilcox_list)
    wilcox_dt[,p_adj := p.adjust(p)]
    wilcox_dt[,p_label := sprintf('%.3g', p_adj)]
    wilcox_dt$y.position = seq(y_plabel[1], y_plabel[2],
                               length.out=nrow(wilcox_dt))
    
    comparisons = lapply(level_vec[level_vec!='iDomain'],
                         function(x){return(c('iDomain', x))})
    beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=0.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_pvalue_manual(wilcox_dt, label = "p_label")
  } else {
    wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                by='group']
    wilcox_dt[,p_adj:=p.adjust(p_value)]
    wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                symbols=symnum$symbols)]
    beeswarm = ggplot(indel_selection,
                      aes_string(x='group', y=y_name, color=group_by)) +
      geom_quasirandom(dodge.width=1) +
      scale_color_manual(values=color) +
      stat_summary(aes_string(group=group_by), fun.=median,
                   fun.min = median, fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red',
                   position= position_dodge(width =1)) +
      geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                inherit.aes=F)
  }
  
  beeswarm = beeswarm +
    theme_bw() +
    geom_vline(xintercept=seq(1.5, length(unique(indel.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank())
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(indel.dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme_bw() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, beeswarm, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}

quantiles_95 <- function(x) {
  r <- quantile(x, probs=c(0.05, 0.25, 0.5, 0.75, 0.95))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

plot_comb_grid_bothplot <- function(bix.dt, opt_df, domains, y_name, min_count=0,
                                   max_count=Inf, group_by=NULL, color=NULL){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(bix.dt[,..in_vec])
    out_sum = rowSums(bix.dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    bix.dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(bix.dt[,table(group)])
  } else {
    group_count_df = data.frame(bix.dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = bix.dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]

  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    theme_bw() +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color=FALSE)  +
    geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", " "))
  
  if (is.null(group_by)){
    boxpl = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=0.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_compare_means(aes(label = ..p.signif..),
                         symnum.args = symnum,
                         method = 'wilcox.test',
                         ref.group='iDomain')
  } else {
    
    boxpl = ggplot(indel_selection,
                   aes_string(x='group', y=y_name, color=group_by)) +
      guides(fill=F) +
      # scale_fill_manual(values=color) +
      scale_color_manual(values=color) +
      # geom_violin(position=position_dodge(width=1)) +
      geom_quasirandom(dodge.width = 1) +
      # stat_summary(fun.data = quantiles_95,
      #              geom="boxplot",
      #              position=position_dodge(width=1),
      #              width = 0.2) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.6,
                   position=position_dodge(width=1))
  }
  boxpl = boxpl +
    theme_bw() +
    geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank()) +
    ylim(0, 1)
  
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(bix.dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme_bw() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(boxpl, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, boxpl, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}
```

# Data import
```{r load data, message=FALSE, warnings=FALSE}
trip_tib_mut_2000 = readRDS(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RDS"))
trip_tib_2000 = readRDS(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RDS"))

TRIP_ChIP <- readRDS("/DATA/usr/x.vergara/XV_ChIPIndels/XV20200902_DDR_RS_revision/XV20200902_DDR_RS_revision/data/xv20200915_DDR_data_CHIP.rds") %>% 
  as_tibble() %>% filter(!antibody %in% c("gH2AX", "LigIII"))
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64" & cell_line == "RSTP2_2000")

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64" & cell_line == "RSTP2_2000")

mean_trip_tib_2000 <- trip_tib_2000 %>%
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

trip_tib$spin_state <- factor(trip_tib$spin_state, levels=c("Speckle",
                                                            "Interior_Act1",
                                                            "Interior_Act2",
                                                            "Interior_Act3",
                                                            "Interior_Repr1",
                                                            "Interior_Repr2",
                                                            "Near_Lm1",
                                                            "Near_Lm2",
                                                            "Near_Lm3",
                                                            "Lamina",
                                                            NA ))

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```

# Main Figure 4

## Panel 4A
```{r Fig4A pathway contribution, fig.height=4, fig.width=8}
pathway_distribution <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios))

barcode_order <- pathway_distribution %>%
  filter(pathway == "NHEJ")  %>%
  arrange(dplyr::desc(ratios)) %>%
  dplyr::select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("SSTR", "MMEJ", "other_indels",  "NHEJ"))

ggplot(pathway_distribution,
       aes(barcode, ratios, width=1)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  xlab("IPR") +
  scale_fill_manual("legend", values = colore) +
  ggtitle("Pathway contribution per barcode") 



### Ordered on efficiency with efficiency plot on top.
# 
pathway_distribution <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64") %>%
  gather(NHEJ, MMEJ, other_indels,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios))

barcode_order <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>%
  dplyr::select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("MMEJ", "other_indels",  "NHEJ"))

trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>% ggplot(.,
       aes(factor(barcode,levels=barcode_order$barcode), efficiency, width=1)) +
  geom_bar(stat = "identity") +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("TIF") +
  xlab("IPR")

ggplot(pathway_distribution,
       aes(barcode, ratios, width=1)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  xlab("IPR") +
  scale_fill_manual("legend", values = colore) +
  ggtitle("Pathway contribution per barcode")


```

### Figure legend

## Panel 4B

```{r Fig4B pathways vs chrom, fig.height=8, fig.width=9}
chrom.mods.cols <- seq(grep("binsize", names(trip_tib))+1, grep("H3K27me3_domain", names(trip_tib))-1)

NHEJ.hist.corr <- CorrelationDFMean(trip_tib, "mean_RSTP2_2000_LBR2_DMSO_64", "MMEJ_MMEJNHEJ", targets = chrom.mods.cols) %>%
  arrange(corr) %>% pull(rowname) %>% factor(.,levels = .)

lims = 1

CorrelationPlotMean(trip_tib, "mean_RSTP2_2000_LBR2_DMSO_64", c("MMEJ_MMEJNHEJ") , targets = NHEJ.hist.corr)
```
### Figure legend
Pearson’s correlation coefficients of the local intensities of 24 chromatin features versus the proportion of +1 (red), -7 (blue) and other indels (black). 

## Panel 4C
```{r Fig4C ratio and pathway distribution, fig.height=3, fig.width=4}
trip_tib%>%
  ggplot(., aes(MMEJ_MMEJNHEJ)) +
  geom_density(size = 1.5) +
  theme_bw(base_size = 16) +
  scale_x_continuous(breaks=seq(0,.7, .1))

trip_tib %>% pull(barcode) %>% unique()  %>% length()

# trip_tib %>% gather(other_indels, NHEJ, MMEJ,
#                     key = "pathway",
#                     value = 'ratios') %>%
#   ggplot(., aes(ratios, fill = pathway, alpha = 0.8)) +
#   geom_density() +
#   theme_bw(base_size = 16) +
#   scale_fill_manual("legend", values = colore)
```
### Figure legend
Distribution of log2(MMEJ/(MMEJ+NHEJ) ratios across all IPRs. 


## Panel 4D
```{r Fig4D MMEJ score vs eff, fig.height=4, fig.width=4}
ggplot(trip_tib, aes(efficiency, MMEJ_MMEJNHEJ)) +
  geom_point() +
  geom_smooth(method= "glm") + 
  theme_bw(base_size = 16) + 
  stat_cor(method = "spearman", label.x.npc = "left", label.y.npc = "bottom")  
```
### Figure legend

<!-- "****" = < 0.0001
"***" = < 0.001
"**" = < 0.01
"*" = < 0.05 -->

## Panel 4E
```{r Fig4E MMEJ ratio per heterochomatin domain, fig.height=6, fig.width=6}
tib.dt = data.table(trip_tib)

domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

indel.dt = unique(tib.dt[exp_pool =="mean_RSTP2_2000_LBR2_DMSO_64",
                         c('exp_pool', 'barcode', 'MMEJ_MMEJNHEJ',
                          domains), with=F])

indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)
#
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_64'], opt_df,
               domain_levels, 'MMEJ_MMEJNHEJ', min_count = 20, y_plabel=c(0.55, 0.70))
```

### Figure legend

## Panel 4F
```{r Fig4F MMEJ balance and heterochromatin signal, fig.height=7, fig.width=12}
trip_tib_triple = trip_tib %>% filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_64", 
                    late_replicating_domain  == "Domain",
                    LAD_domain == "Domain",
                    H3K9me2_domain == "Domain",
                    H3K27me3_domain == "iDomain") %>% 
  distinct(barcode, MMEJ_MMEJNHEJ, H3K9me2, LMNB1, late_replicating)

p1 = trip_tib_triple %>%
  distinct(barcode, H3K9me2, MMEJ_MMEJNHEJ) %>%
  pivot_longer(!c(barcode, MMEJ_MMEJNHEJ), names_to = "modification", values_to = "value") %>%
  ggplot(., aes(modification, MMEJ_MMEJNHEJ, color = value))  +
  geom_quasirandom(size = 3) +
  theme_bw(base_size = 16) + 
  stat_cor(aes(MMEJ_MMEJNHEJ, value), method = "spearman", label.y = .55, label.x = 1.2)  +
  ylab("MMEJ/(MMEJ+NHEJ)") +
  theme(legend.position = "right") + 
  scale_color_viridis()

p2 = trip_tib_triple %>%
  distinct(barcode, LMNB1, MMEJ_MMEJNHEJ) %>%
  pivot_longer(!c(barcode, MMEJ_MMEJNHEJ), names_to = "modification", values_to = "value") %>%
  ggplot(., aes(modification, MMEJ_MMEJNHEJ, color = value))  +
  geom_quasirandom(size = 3) +
  theme_bw(base_size = 16) + 
  stat_cor(aes(MMEJ_MMEJNHEJ, value), method = "spearman", label.y = .55, label.x = 1.2)  +
  ylab("MMEJ/(MMEJ+NHEJ)") +
  theme(legend.position = "right") + scale_color_viridis()

p3 = trip_tib_triple %>%
  distinct(barcode, late_replicating, MMEJ_MMEJNHEJ) %>%
  pivot_longer(!c(barcode, MMEJ_MMEJNHEJ), names_to = "modification", values_to = "value") %>%
  ggplot(., aes(modification, MMEJ_MMEJNHEJ, color = value))  +
  geom_quasirandom(size = 3) +
  theme_bw(base_size = 16) + 
  stat_cor(aes(MMEJ_MMEJNHEJ, value), method = "spearman", label.y = .55, label.x = 1.2)  +
  ylab("MMEJ/(MMEJ+NHEJ)") +
  theme(legend.position = "right") + 
  scale_color_viridis()

plot_grid(p1, p2, p3, nrow = 1)
```


####### XABI CHIP       
## Panel 4G

```{r Fig4G DNA damage response protein pull down, fig.height=4, fig.width=10}
mean_trip_chip <- TRIP_ChIP %>%
         dplyr::group_by(antibody, plasmid, barcode, chromatin) %>%
         dplyr::summarise(norm.pull.down = mean(norm.pull.down),
                          replicate = "mean")

chrom_colors2 = c( c("Euchromatin" = "#F7941D", "other-heterochromatin" = "#838687",  "H3K27me3" = "#D21F8A", "Triple Heterochromatin" = "#662D91"))

ggplot(mean_trip_chip, aes(plasmid,log2(norm.pull.down), color = chromatin)) +
  geom_quasirandom() +
  facet_grid(~ antibody) +
  coord_cartesian() +
  theme_bw(base_size = 16) + 
  scale_color_manual(values = chrom_colors2) +
  ylim(-4.5, 2) +
  ylab("log2(Normalized pull down counts)") +
  ggtitle("DNA damage response protein pull down") +
  theme(legend.position = "bottom")
```


### Figure legend


## Panel 4H

```{r Fig4H normalized ChIP vs adjusted efficiency, fig.height=4, fig.width=10}
## Prepare models for LBR1, LBR12 & LBR15 to use to "spread" the LBR2 efficiency. This will allow us to maybe bipass the saturation in highly efficient IPRs.
trip_tib_otherguides <- trip_tib_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR1_DMSO_64",
                         "mean_RSTP2_2000_LBR2_DMSO_64",
                         "mean_RSTP2_2000_LBR12_DMSO_64",
                         "mean_RSTP2_2000_LBR15_DMSO_64"))
otherguides_barcodes <- trip_tib_otherguides %>% 
  group_by(barcode) %>% 
  dplyr::summarise(count = n()) %>% 
  filter(count > 1) %>% 
  pull(barcode)

length(otherguides_barcodes)

clone5_barcodes <- trip_tib_2000 %>% filter(cell_line == "RSTP2_clone5") %>% distinct(barcode) %>% pull()

trip_tib_otherguides$plasmid <- factor(trip_tib_otherguides$plasmid, levels = c("LBR1", "LBR12", "LBR15" ,"LBR2"))

trip_tib_otherguides_wide <- trip_tib_otherguides %>% filter(barcode %in% otherguides_barcodes) %>% dplyr::select(barcode, plasmid, efficiency) %>% spread(plasmid, efficiency) 


# trip_tib_otherguides_wide %>% ggplot(., aes(LBR2, LBR15))  +  
#   geom_point() +
#   geom_smooth(method= "loess") +
#   ylim(0, 1) +
#   xlim(0, 1) +
#   theme_bw(base_size = 16) +
#   stat_cor(method = "spearman") + 
#   coord_fixed()



# Make the loess model on LBR1, 12 & 15 with LBR2. 
loess_LBR1 <- loess(LBR1 ~ LBR2, trip_tib_otherguides_wide, 
                    control = loess.control(surface = "direct")) # This option allows for extrapolation
loess_LBR12 <- loess(LBR12 ~ LBR2, trip_tib_otherguides_wide, 
                     control = loess.control(surface = "direct"))
loess_LBR15 <- loess(LBR15 ~ LBR2, trip_tib_otherguides_wide, 
                     control = loess.control(surface = "direct"))


fc.mean.DDR.model <- TRIP_ChIP %>% 
  left_join(TRIP_ChIP %>% 
              dplyr::group_by(chromatin,antibody,barcode,plasmid) %>% 
              dplyr::summarise(m.fc = mean(fold.change), 
                               m.TIF = mean(freqcut), 
                               m.MMEJ = mean(MMEJscore),
                               m.pulldown = mean(norm.pull.down), 
                               by = c("chromatin","antibody","barcode","plasmid"))) %>% 
  filter(!antibody %in% c("LigIII","gH2AX")) %>% 
  distinct(barcode, antibody, m.fc, m.TIF, 
           m.pulldown, m.MMEJ, plasmid, chromatin) %>%
  ungroup() %>%
  pivot_wider(names_from = plasmid, 
              values_from = c(m.fc, m.pulldown, m.MMEJ, m.TIF)) %>% 
  mutate(norm.pulldown = m.pulldown_LBR2 - m.pulldown_EV) %>%
  pivot_longer(-c(antibody, barcode, chromatin, norm.pulldown),
               names_to = c(".value", "plasmid"),
               names_sep = "_")

fc.mean.DDR.model = fc.mean.DDR.model %>% 
  mutate(LBR1_pred = predict(loess_LBR1, m.TIF, se = TRUE)$fit, 
         LBR12_pred = predict(loess_LBR12, m.TIF, se = TRUE)$fit,
         LBR15_pred = predict(loess_LBR15, m.TIF, se = TRUE)$fit) %>% 
  mutate(mean_pred = rowMeans(dplyr::select(., c("LBR1_pred", "LBR12_pred", "LBR15_pred"))),
         log2_mean_pred = log2(mean_pred),
         log2_efficiency = log2(m.TIF),
         log2_m.pulldown = log2(norm.pulldown))

### Modeling with decompressed efficiency (based on LBR1, LBR12 & LBR15)
lm_guides_LigIV <- nls(log2_m.pulldown ~ a + b*log2_mean_pred, algorithm = "port", 
                       start = c(a = 0, b = 1), lower = c(a = -Inf,b = 1), 
                       upper = c(Inf, b = 1), 
                       data = filter(fc.mean.DDR.model,
                                     plasmid == "LBR2", antibody == "LigIV"))
lm_guides_MRE11 <- nls(log2_m.pulldown ~ a + b*log2_mean_pred, algorithm = "port", 
                       start = c(a = 0, b = 1), lower = c(a = -Inf,b = 1), 
                       upper = c(Inf, b = 1), 
                       data = filter(fc.mean.DDR.model,
                                     plasmid == "LBR2", antibody == "MRE11"))
lm_guides_POLQ <- nls(log2_m.pulldown ~ a + b*log2_mean_pred, algorithm = "port", 
                      start = c(a = 0, b = 1), lower = c(a = -Inf,b = 1), 
                      upper = c(Inf, b = 1), 
                      data = filter(fc.mean.DDR.model,
                                    plasmid == "LBR2", antibody == "POLQ"))
lm_guides_Rad51 <- nls(log2_m.pulldown ~ a + b*log2_mean_pred, algorithm = "port", 
                       start = c(a = 0, b = 1), lower = c(a = -Inf,b = 1), 
                       upper = c(Inf, b = 1), 
                       data = filter(fc.mean.DDR.model,
                                     plasmid == "LBR2", antibody == "Rad51"))

antibody = c("LigIV", "MRE11", "POLQ", "Rad51")
intercepts = c(coef(lm_guides_LigIV)[[1]], 
               coef(lm_guides_MRE11)[[1]], 
               coef(lm_guides_POLQ)[[1]], 
               coef(lm_guides_Rad51)[[1]])
intercept_df = data.frame(antibody, intercepts)


## Plotting the decompressed total indel frequency, with in red the predicted relationshop (model with slope = 1)
# fc.mean.DDR.model %>% 
#   filter(plasmid == "LBR2") %>% 
#   ggplot(., aes(log2_mean_pred, log2_m.pulldown, color = chromatin, group = 1)) + 
#   geom_point(size = 3) +
#   geom_smooth(method = lm) + 
#   geom_abline(data= intercept_df, aes(intercept = intercepts, slope = 1), size = 1, color = "red") +
#   stat_cor() + 
#   facet_grid(~ antibody) + 
#   scale_color_manual(values = chrom_colors2) +
#   theme_bw(base_size = 16) + 
#   ylab("log2(mean pull down)") + 
#   xlab("log2(average modelled gene editing frequency)") + 
#   ggtitle("modelled gene editing frequency vs. mean pull down") +
#   theme(legend.position = "bottom")
```

### Figure legend

## Panel 4I
```{r Fig4I residuals vs MMEJ:NHEJ balance, fig.height=4, fig.width=10}
# Calculate the residuals from the datapoints in the previous figure from the predicted relationship
LigIV_res <- lm_guides_LigIV %>% augment %>% mutate(antibody = "LigIV")
MRE11_res <- lm_guides_MRE11 %>% augment  %>% mutate(antibody = "MRE11")
POLQ_res <- lm_guides_POLQ %>% augment %>% mutate(antibody = "POLQ")
Rad51_res <- lm_guides_Rad51 %>% augment %>% mutate(antibody = "Rad51")

resids_model = rbind(LigIV_res, MRE11_res, POLQ_res, Rad51_res)

# fc.mean.DDR.model %>% filter(plasmid == "LBR2") %>% left_join(resids_model) %>%
#   ggplot(., aes(m.TIF, .resid, color = chromatin)) + geom_jitter(width = .2, size = 3) +facet_wrap(. ~ antibody) + 
#     scale_color_manual(values = chrom_colors2) +
#     theme_bw(base_size = 16) + 
#     ylab("residuals") + 
#     xlab("modelled total indel frequency") + 
#     ggtitle("residuals from the red model (slope = 1) per antibody and chromatin group")

## Plot the residuals against the MMEJ:NHEJ balance. 
fc.mean.DDR.model %>% 
  filter(plasmid == "LBR2") %>% 
  left_join(resids_model) %>%
  ggplot(., aes(m.MMEJ, .resid, color = chromatin, group = 1)) + geom_point(size = 3) +
  facet_grid(. ~ antibody) + 
  scale_color_manual(values = chrom_colors2) +
  theme_bw(base_size = 16) + 
  geom_smooth(method = lm) + 
  stat_cor(method = "pearson") + 
  ylab("residuals") + 
  xlab("MMEJ:NHEJ balance") + 
  ggtitle("residuals from the red model (slope = 1) vs MMEJ:NHEJ balance per antibody") +
  theme(legend.position = "bottom")

fc.mean.DDR.model %>% 
  filter(plasmid == "LBR2", chromatin %in% c("Euchromatin", "Triple Heterochromatin")) %>% 
  mutate(chromatin = ifelse(chromatin == "Euchromatin", "euchrom", "triple")) %>%
  left_join(resids_model, by = c("antibody", "log2_mean_pred", "log2_m.pulldown")) %>% 
  distinct(antibody, chromatin, .resid, barcode) %>% 
  nest(data = - antibody) %>% 
  mutate(
    test = map(data, ~ wilcox.test(.$.resid ~ .$chromatin)), # S3 list-col
    tidied = map(test, tidy)
  ) %>% 
  unnest(tidied)

fc.mean.DDR.model %>% 
  filter(plasmid == "LBR2", chromatin %in% c("Euchromatin", "Triple Heterochromatin")) %>% 
  mutate(chromatin = ifelse(chromatin == "Euchromatin", "euchrom", "triple")) %>%
  left_join(resids_model, by = c("antibody", "log2_mean_pred", "log2_m.pulldown")) %>% 
  distinct(antibody, chromatin, .resid, barcode) %>%
  ggplot(aes(chromatin, .resid)) + 
  geom_quasirandom() + 
  theme_bw(base_size = 16) +
  facet_grid(. ~ antibody) +
  stat_compare_means() +
  stat_summary(fun=mean, geom="crossbar", color='red')
```



# Supplementary Figure 4
## Panel S4A
```{r FigS4A scatterplot pathways vs lmnb1 and K4me1, fig.height=8, fig.width=12}
trip_tib %>%
  dplyr::select(barcode, MMEJ, NHEJ, other_indels, LMNB1, H3K4me1) %>%
  group_by(barcode) %>%
  gather(MMEJ, NHEJ, other_indels, key = "pathway", value = "ratio") %>%
  gather(LMNB1, H3K4me1, key = "feature", value = "value") %>%
  ggplot(., aes(ratio, value))  +  
  geom_point() +
  geom_smooth(method= "lm") +
  theme_bw(base_size = 16) +
  stat_cor(method = "spearman", label.x = .25, label.y = .7) +
  facet_wrap(feature ~ pathway, scales = "free", ncol = 3)
```

### Figure legend

## Panel S4B
```{r FigS4B MMEJ score vs eff, fig.height=3, fig.width=4}
trip_tib %>% mutate(cell_pool = gsub(".*[.]", "", barcode)) %>%
  ggplot(., aes(MMEJ_MMEJNHEJ, color = cell_pool)) +
  geom_density(size = 1.5) +
  theme_bw(base_size = 16) +
  scale_x_continuous(breaks=seq(0,.7, .1))
```
### Figure legend

<!-- "****" = < 0.0001
"***" = < 0.001
"**" = < 0.01
"*" = < 0.05 -->

## Panel S4C
```{r FigS4C MMEJ ratio per heterochomatin domain supp, fig.height=7, fig.width=14}
tib.dt = data.table(trip_tib)

domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

indel.dt = unique(tib.dt[exp_pool =="mean_RSTP2_2000_LBR2_DMSO_64",
                         c('exp_pool', 'barcode', 'MMEJ_MMEJNHEJ',
                           domains), with=F])

indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)
#
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_64'], opt_df,
               domain_levels, 'MMEJ_MMEJNHEJ', y_plabel=c(0.55, 0.70))
```

### Figure legend

## Panel S4D


```{r FigS4D MMEJ balance other guides, fig.height=6, fig.width=10}
MMEJbal_otherguides_tib = readRDS("/DATA/projects/DSBrepair/data/R/rs20201005_clone5_otherguides_MMEJbalance_chromdomains.RDS") 

MMEJbal_otherguides_het_eu_tib = MMEJbal_otherguides_tib %>% 
  mutate(chromatin = ifelse(chrom_grp == "euchromatin", "euchromatin", "heterochromatin")) %>%
  filter(drug == "DMSO", !is.na(chromatin))

MMEJbal_otherguides_het_eu_tib %>%
  ggplot(aes(chromatin, MMEJ_MMEJNHEJ)) + 
  geom_quasirandom() + 
  theme_bw(base_size = 16) +
  facet_grid(. ~ plasmid) +
  stat_compare_means() +
  stat_summary(fun=mean, geom="crossbar", color='red')


MMEJbal_otherguides_triple_tib = MMEJbal_otherguides_tib %>% 
  mutate(LAD_domain == ifelse(is.na(LAD_domain), "iDomain", LAD_domain)) %>%
  filter(chrom_grp %in% c("triple", "euchromatin"), 
         drug == "DMSO")

MMEJbal_otherguides_triple_tib %>%
  ggplot(aes(chrom_grp, MMEJ_MMEJNHEJ)) + 
  geom_quasirandom() + 
  theme_bw(base_size = 16) +
  facet_grid(. ~ plasmid) +
  stat_compare_means() +
  stat_summary(fun=mean, geom="crossbar", color='red')
```
## Panel 4E

```{r FigS4E MMEJ balance vs. mean fold change, fig.height=4, fig.width=10}
fc.mean.DDR.model %>% 
  filter(plasmid == "LBR2") %>%
  ggplot(., aes(m.MMEJ,log2(m.fc), color = chromatin, group = 1)) + 
  geom_point(size = 3) +
  geom_smooth(method = lm) + 
  scale_color_manual(values = chrom_colors2) +
  stat_cor() + 
  facet_grid(. ~ antibody) + 
  theme_bw(base_size = 16) + 
  coord_cartesian() +
  ylab("log2(average protein ChIP cut/uncut) at 16h") + 
  xlab("MMEJ/(MMMEJ+NHEJ) at 72h") + 
  ggtitle("MMEJ:NHEJ balance vs. mean fold change") +
  theme(legend.position = "bottom")
```

### Figure legend

# Bibliography
```{r citations}
cite_packages()
```


# Session Info
```{r}
paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```
