---
title: "Figure 7 and supplementary 7-8"
author: "Ruben Schep"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    theme: journal #cerulean
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
  editor_options:
    chunk_output_type: console
---

# knitr document van Steensel lab
```{r}
StartTime <-Sys.time()
library(knitr)

# 8-digit Date tag:
Date <- substr(gsub("-","",Sys.time()),1,8)

## Select outdir
out.dir = paste0("figures/rs", Date, "/")
dir.create(out.dir)
opts_chunk$set(dev=c('png', 'pdf'), fig.path = file.path(out.dir))
pdf.options(useDingbats = FALSE)
```

# Introduction

## Description of Data

# Data processing
## Path, Libraries, Parameters and Useful Functions
```{r setup, message=FALSE, warnings=FALSE}
# INPUT DIRECTORY
in.dir.date = 20201029
in.dir = paste0("/DATA/projects/DSBrepair/data/R/rs", in.dir.date, "/")

# libraries:
library(ggplot2)
library(ggpubr)
library(corrr)
library(Hmisc)
library(ggbeeswarm)
library(dplyr)
library(tidyr)
library(tibble)
library(ggpmisc)
library(cowplot)
library(data.table)
library(report)

# library(ggpmisc)
# library(glmnet)
# library(cowplot)
# library(mda)
# library(earth)
# library(yaml)
# library(vip)
# library(caret)
# library(scales)
# library(glmnet)
# library(gridExtra)
# library(ggcorrplot)
# library(bnlearn)
# library(pheatmap)
# library(ppcor)
# library(parallel)


# Color codes used in this script
colore <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colores <- c("wt" = "#808184", "other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")
colory <- c("wt" = "#808184", "other" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "MMEJplusNHEJ" = "#EE8A2A", "all" = "black", "SSTR" = "#007A4C")
measure_color = c(NHEJ='#e1251b', MMEJ='#26419a', SSTR='#007A4C', None='grey')
```

## Custom functions
Functions used thoughout this script.
```{r, message=FALSE, warnings=FALSE}
SetFileName <- function(filename, initials) {
  # Set filename with extension and initials to make filename with date integrated.
  filename <- substitute(filename)
  initials <- substitute(initials)
  filename <- paste0(initials, substr(gsub("-","",Sys.time()),1,8), "_", filename)
  filename
}

CorrelationDFMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  hist.corr.list <- data[data$exp_pool==condition_name, ] %>% dplyr::select(all_of(feature_name), targets) %>% as.matrix() %>% rcorr(type = "pearson")

  histone.corr.df <- data.frame(rowname = names(hist.corr.list$r[1, -1]),
                                corr = hist.corr.list$r[1, -1],
                                pvalue = hist.corr.list$P[1, -1])

  return(histone.corr.df=histone.corr.df)
}


CorrelationPlotMean <- function(data, condition_name, feature_name, targets, p.value=0.001){
  cor.df = data.frame()
  for (i in feature_name){
    corr.list <- data %>% filter(exp_pool==condition_name) %>% dplyr::select(all_of(i), as.character(targets)) %>% as.matrix() %>% rcorr(type = "pearson")

    df <- data.frame(rowname = names(corr.list$r[1, -1]),
                                corr = corr.list$r[1, -1],
                                pvalue = corr.list$P[1, -1],
                                feature = i)
    cor.df <- rbind(cor.df, df)
  }
  p <- cor.df %>%
    transform(rowname = factor(rowname, levels = levels(targets)),
              corr = corr,
              pvalue = pvalue) %>%
    ggplot(aes(x = rowname, y = corr, fill = feature, alpha = pvalue < p.value)) +
    geom_bar(stat = "identity", position = "dodge") +
    ylab("Pathway correlation with chromatin feature") +
    xlab("chromatin feature") +
    theme_bw(base_size = 16) +
    scale_fill_manual("legend", values = colores) +
    scale_y_continuous(breaks=seq(-1,1, .1)) +
    coord_flip()
  return(list(dataframe=cor.df, plot = p))
}

cor_test <- function(y, X, left, right){
    cor_list = lapply(X, function(x){
        cor.test(x,y)
    })
    p_vec = unlist(lapply(cor_list, function(x){x$p.value}))
    cor_vec = unlist(lapply(cor_list, function(x){x$estimate}))
    p_adj = p.adjust(p_vec)
    color = ifelse(p_adj < 0.05, ifelse(cor_vec > 0, left, right),
                   'None')
    importance = abs(cor_vec)
    target=factor(colnames(X), levels=rownames(clustering))
    feature_group = clustering[colnames(X), ]
    return(list(correlation=cor_vec,importance = importance, p=p_vec,
                p_adj=p_adj, target=target, color=color,
                feature_group=feature_group))
}

get_combinations <- function(domain_vec){
    n_domain = length(domain_vec)
    get_ivec <-function (last_x, n_domain){
    	if (last_x < n_domain){
    		i_vec = c(0, c((last_x+1):n_domain))
    	} else{
    		i_vec = c(0)
    	}
    }

    combine <- function(x_new, x_vec, n_domain){
        new_vec = c(x_vec, x_new)
        if (length(new_vec)==n_domain){
            return(new_vec)
        } else {
            i_vec = get_ivec(max(new_vec), n_domain)
            x_list = lapply(i_vec, combine, x_vec=new_vec,
                            n_domain=n_domain)
            return(do.call(rbind, x_list))
        }
    }

    comb_list = lapply(c(0:n_domain), combine, x_vec=c(), n_domain=n_domain)

    comb = do.call(rbind, comb_list)
    comb_df = unique(t(apply(comb, 1, function(x){x[order(x)]})))

    name_list = apply(comb_df, 1, function(x){
    	name = paste(domain_vec[x], collapse='-')
    	if (name == ''){
    		name = 'iDomain'
    	}
    	return(name)
    })

    opt_list = lapply(domain_vec, function(d){
    	 apply(comb_df, 1, function(x){
    		opt = domain_vec[x]
    		return(d%in%opt)
    	})
    })

    opt_df = do.call(cbind, opt_list)
    colnames(opt_df) = domain_vec
    rownames(opt_df) = name_list

    return(opt_df)
}

test_wilcox <- function(dt, group_by='condition', y_name='MMEJ_MMEJNHEJ'){
    bc_vec = dt[,unique(barcode)]
    name1 = unlist(unique(dt[,group_by, with=F])[1])
    dt_c = dt[get(group_by)==eval(name1), ]
    dt_s = dt[get(group_by)!=eval(name1), ]
    setkey(dt_c, 'barcode')
    setkey(dt_s, 'barcode')

    w = wilcox.test(unlist(dt_c[bc_vec,..y_name]),
                    unlist(dt_s[bc_vec,..y_name]),
                    paired=T)
    return(list(p_value = w$p.value))
}

quantiles_95 <- function(x) {
  r <- quantile(x, probs=c(0.05, 0.25, 0.5, 0.75, 0.95))
  names(r) <- c("ymin", "lower", "middle", "upper", "ymax")
  r
}

plot_comb_grid <- function(bix.dt, opt_df, domains, y_name, min_count=0,
                           max_count=Inf, group_by=NULL, color=NULL,
                           y_plabel=c(-3, 3)){

    opt_melt = reshape2::melt(opt_df)
    colnames(opt_melt) = c('group', 'domain', 'is_in')

    opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
    opt_melt$domain = factor(opt_melt$domain, levels=domains)


    group_list = lapply(rownames(opt_df), function(opt_name){
    	opt = opt_df[opt_name, ]
    	in_vec = domains[opt]
    	out_vec = domains[!opt]

    	in_sum = rowSums(bix.dt[,..in_vec])
    	out_sum = rowSums(bix.dt[,..out_vec])
    	if (length(in_vec)==0){
    		in_group = which(out_sum == 0)
    	} else if (length(out_vec) == 0){
    		in_group = which(in_sum==length(in_vec))
    	} else  {
    		in_group = which(in_sum == length(in_vec) & out_sum==0)
    	}
    	return(in_group)
    })

    names(group_list) = rownames(opt_df)
    for (opt_name in names(group_list)){
    	i_vec = group_list[[opt_name]]
    	bix.dt[i_vec, group := opt_name]
    }
    if (is.null(group_by)){
        group_count_df = data.frame(bix.dt[,table(group)])
    } else {
        group_count_df = data.frame(bix.dt[,table(group)/length(unique(get(group_by)))])
    }
    group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
    # print(group_count_df)
    group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
    group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
    group_levels =  group_count_sel$group
    opt_sel = opt_melt[opt_melt$group %in% group_levels, ]

    opt_sel$group = factor(opt_sel$group, levels=group_levels)
    # print(opt_sel)

    indel_selection = bix.dt[group%in%group_levels, ]
    indel_selection[,group:=factor(group, levels=group_levels)]


    group_count_sel$group = factor(group_count_sel$group, levels=group_levels)

    count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
                            geom_label() +
                            theme_void()



    dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    	geom_point(size=2) +
    	geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    	theme_bw() +
    	scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    	guides(color=FALSE)  +
    	geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
    			   lwd=0.5, colour="grey90") +
    	theme(axis.title=element_blank(),
    		  axis.text=element_blank(),
    		  axis.ticks=element_blank(),
    		  panel.grid.major = element_blank())

    symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                  symbols = c("****", "***", "**", "*", " "))


    if (is.null(group_by)){
      level_vec = indel_selection[, levels(group)]
      wilcox_list = lapply(level_vec[level_vec!='iDomain'],
                           function(g){
                               x = indel_selection[group==g, y_name, with=F]
                               y = indel_selection[group=='iDomain', y_name,
                                                 with=F]
                               w = wilcox.test(unlist(x),unlist(y))
                               return(data.table('group1'='iDomain', 'group2'=g,
                                                 'p'=w$p.value))
                           })
      wilcox_dt = do.call(rbind, wilcox_list)
      wilcox_dt[,p_adj := p.adjust(p)]
      wilcox_dt[,p_label := sprintf('%.3g', p_adj)]
      wilcox_dt$y.position = seq(y_plabel[1], y_plabel[2],
                                 length.out=nrow(wilcox_dt))

      comparisons = lapply(level_vec[level_vec!='iDomain'],
                           function(x){return(c('iDomain', x))})
      beeswarm = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
        geom_quasirandom(width=0.3) +
        stat_summary(fun=median, fun.min = median,
                     fun.max = median,
                     geom="crossbar", width = 0.5,
                     color='red') +
        stat_pvalue_manual(wilcox_dt, label = "p_label")
    } else {
        wilcox_dt = indel_selection[,test_wilcox(.SD, group_by, y_name),
                                    by='group']
        wilcox_dt[,p_adj:=p.adjust(p_value)]
        wilcox_dt[,p_signif:=symnum(p_adj, cutpoints=symnum$cutpoints,
                                    symbols=symnum$symbols)]
        beeswarm = ggplot(indel_selection,
                          aes_string(x='group', y=y_name, color=group_by)) +
        			   geom_quasirandom(dodge.width=1) +
                       scale_color_manual(values=color) +
                       stat_summary(aes_string(group=group_by), fun=median,
                                    fun.min = median, fun.max = median,
                                    geom="crossbar", width = 0.5,
                                    color='red',
                                    position= position_dodge(width =1)) +
                       geom_text(data=wilcox_dt, aes(x=group, y=0.8, label=p_signif),
                                 inherit.aes=F)
    }
    beeswarm = beeswarm +
    			  theme_bw() +
    			  geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
    			             lwd=0.5, colour="grey90") +
    			  theme(axis.title=element_blank(),
    			        axis.text.x=element_blank(),
      				    panel.grid.major.x = element_blank(),
    				    axis.ticks.x=element_blank())


    count = data.frame(domain=gsub('_domain', '', domains),
    	               count=colSums(bix.dt[,..domains], na.rm=T))

    count$domain = factor(count$domain, levels=domain_levels)


    histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    				geom_label() +
    				theme_bw() +
    				theme(plot.background = element_blank(),
        				  panel.grid.major = element_blank(),
        				  panel.grid.minor = element_blank(),
        				  panel.border = element_blank(),
    					  axis.title=element_blank(),
    	  			      axis.text.x=element_blank(),
    	  				  axis.ticks=element_blank())

    empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
                geom_point() +
                theme_void()

    count_beeswarm = plot_grid(beeswarm, count_plot, ncol=1, rel_heights=c(20,1),
                               align='v', axis='lr')
    return(plot_grid(empty, beeswarm, empty, count_plot,
                     histogram, dots, nrow=3,
                     rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                     axis='tlbr'))
}


plot_comb_grid_boxplot <- function(bix.dt, opt_df, domains, y_name, min_count=0,
                                   max_count=Inf, group_by=NULL, color=NULL){
  
  opt_melt = reshape2::melt(opt_df)
  colnames(opt_melt) = c('group', 'domain', 'is_in')
  
  opt_melt$is_in = ifelse(opt_melt$is_in, 'in', 'out')
  opt_melt$domain = factor(opt_melt$domain, levels=domains)
  
  
  group_list = lapply(rownames(opt_df), function(opt_name){
    opt = opt_df[opt_name, ]
    in_vec = domains[opt]
    out_vec = domains[!opt]
    
    in_sum = rowSums(bix.dt[,..in_vec])
    out_sum = rowSums(bix.dt[,..out_vec])
    if (length(in_vec)==0){
      in_group = which(out_sum == 0)
    } else if (length(out_vec) == 0){
      in_group = which(in_sum==length(in_vec))
    } else  {
      in_group = which(in_sum == length(in_vec) & out_sum==0)
    }
    return(in_group)
  })
  
  names(group_list) = rownames(opt_df)
  for (opt_name in names(group_list)){
    i_vec = group_list[[opt_name]]
    bix.dt[i_vec, group := opt_name]
  }
  if (is.null(group_by)){
    group_count_df = data.frame(bix.dt[,table(group)])
  } else {
    group_count_df = data.frame(bix.dt[,table(group)/length(unique(get(group_by)))])
  }
  group_count_df = group_count_df[order(match(group_count_df$group, rownames(opt_df))),]
  # print(group_count_df)
  group_count_sel = group_count_df[group_count_df$Freq>=min_count,]
  group_count_sel = group_count_sel[group_count_sel$Freq<=max_count,]
  group_levels =  group_count_sel$group
  opt_sel = opt_melt[opt_melt$group %in% group_levels, ]
  
  opt_sel$group = factor(opt_sel$group, levels=group_levels)
  # print(opt_sel)
  
  indel_selection = bix.dt[group%in%group_levels, ]
  indel_selection[,group:=factor(group, levels=group_levels)]
  
  group_count_sel$group = factor(group_count_sel$group, levels=group_levels)
  
  count_plot = ggplot(group_count_sel, aes(x=group,y=1, label=Freq)) +
    geom_label() +
    theme_void()
  
  
  
  dots = ggplot(opt_sel, aes(x=group, y=domain, color=is_in)) +
    geom_point(size=2) +
    geom_line(data=opt_sel[opt_sel$is_in=='in',], aes(group=group)) +
    theme_bw() +
    scale_color_manual(values=c('in'='black', 'out'='grey70')) +
    guides(color=FALSE)  +
    geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text=element_blank(),
          axis.ticks=element_blank(),
          panel.grid.major = element_blank())
  
  symnum = list(cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 1),
                symbols = c("****", "***", "**", "*", " "))

  if (is.null(group_by)){
    boxpl = ggplot(indel_selection, aes_string(x='group', y=y_name)) +
      geom_quasirandom(width=.3) +
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red') +
      stat_compare_means(aes(label = ..p.signif..),
                         symnum.args = symnum,
                         method = 'wilcox.test',
                         ref.group='iDomain')
  } else {
    
    boxpl = ggplot(indel_selection,
                   aes_string(x='group', y=y_name, color=group_by, fill = group_by)) +
      geom_quasirandom(width=0.2, dodge.width = .9) + 
      stat_summary(fun=median, fun.min = median,
                   fun.max = median,
                   geom="crossbar", width = 0.5,
                   color='red', position=position_dodge(width=.9)) +
      guides(fill=F) +
      scale_fill_manual(values=color) +
      scale_color_manual(values=color)
      # stat_summary(fun.data = quantiles_95, 
      #              geom="boxplot", 
      #              position=position_dodge(width=1))
  }
  boxpl = boxpl +
    theme_bw() +
    geom_vline(xintercept=seq(1.5, length(unique(bix.dt$group))-0.5, 1),
               lwd=0.5, colour="grey90") +
    theme(axis.title=element_blank(),
          axis.text.x=element_blank(),
          panel.grid.major.x = element_blank(),
          axis.ticks.x=element_blank())
  
  
  count = data.frame(domain=gsub('_domain', '', domains),
                     count=colSums(bix.dt[,..domains], na.rm=T))
  
  count$domain = factor(count$domain, levels=domain_levels)
  
  
  histogram = ggplot(count, aes(x=1, y=domain, label = count)) +
    geom_label() +
    theme_bw() +
    theme(plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.title=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks=element_blank())
  
  empty = ggplot(data=data.frame(1,1),aes(x=1,y=1)) +
    geom_point() +
    theme_void()
  
  count_beeswarm = plot_grid(boxpl, count_plot, ncol=1, rel_heights=c(20,1),
                             align='v', axis='lr')
  return(plot_grid(empty, boxpl, empty, count_plot,
                   histogram, dots, nrow=3,
                   rel_widths=c(1,10), rel_heights=c(15,1,5), align='vh',
                   axis='tlbr'))
}

```

# Data import
```{r data loading, message=FALSE, warnings=FALSE}
trip_tib_mut_2000 = readRDS(paste0(in.dir, "RSTP2_Indel_Chromatin_2kb.RDS"))
trip_tib_2000 = readRDS(paste0(in.dir, "RSTP2_IndelRatios_Chromatin_2kb.RDS"))
```


## Data loading

```{r preprocess experiments}
trip_tib_mut <- copy(trip_tib_2000)
trip_tib_2000 <- trip_tib_2000 %>% distinct(barcode, exp, .keep_all = TRUE)

trip_tib <- trip_tib_2000 %>%
  filter(exp_pool  %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_NU7441_64", "mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

trip_tib_mut <- trip_tib_mut_2000 %>%
  filter(exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_64", "mean_RSTP2_2000_LBR2_NU7441_64", "mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

mean_trip_tib_2000 <- trip_tib_2000 %>%
  gather(NHEJ, MMEJ, other_indels, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

mean_trip_tib_2000$pathway <- factor(mean_trip_tib_2000$pathway, levels=c("other_indels", "MMEJ", "NHEJ"))
mean_trip_tib_2000 <- mean_trip_tib_2000[!is.na(mean_trip_tib_2000$ratios), ]
```

# Main Figure 7

## Panel 7A

```{r Fig7A pathways vs contribution ssODN, fig.height=6, fig.width=5}
trip_tib_pathways <- trip_tib %>%
  gather(NHEJ, MMEJ, other_indels, SSTR, key = "pathway", value = 'ratios') %>%
  distinct(barcode, exp, pathway, .keep_all = TRUE)

exp_order <- c("mean_RSTP2_2000_LBR2_DMSO_64",
              "mean_RSTP2_2000_LBR2_NU7441_64",
              "mean_RSTP2_2000_LBR2_DMSO_ssODN_64",
              "mean_RSTP2_2000_LBR2_NU7441_ssODN_64")

trip_tib_pathways$pathway <- factor(trip_tib_pathways$pathway, levels=c("other_indels", "MMEJ", "SSTR", "NHEJ"))
trip_tib_pathways$exp_pool <- factor(trip_tib_pathways$exp_pool, levels=exp_order)
trip_tib_pathways <- trip_tib_pathways[!is.na(trip_tib_pathways$ratios), ]

pathway_contribution <- trip_tib_pathways %>% group_by(exp_pool, pathway) %>% dplyr::summarise(pathway_contr = mean(ratios))

colores <- c("other_indels" = "black", "NHEJ" = "#E1251B", "MMEJ" = "#223DA0", "SSTR" = "#007A4C")

p <- ggplot(pathway_contribution, aes(exp_pool, pathway_contr, fill = pathway, order = pathway)) +
  geom_bar(stat = "identity", position = "fill")  +
  theme_bw(base_size = 16)
p + scale_fill_manual("legend", values = colores) +
  scale_x_discrete(labels=c("mean_RSTP2_2000_LBR2_DMSO_64" = "DMSO",
                            "mean_RSTP2_2000_LBR2_NU7441_64" = "ssODN + DMSO",
                            "mean_RSTP2_2000_LBR2_DMSO_ssODN_64" = "NU7441",
                            "mean_RSTP2_2000_LBR2_NU7441_ssODN_64" = "ssODN + NU7441")) +
  labs(x = element_blank(),
       y = "abundance")
```

### Figure legend
(A) Average pathway contribution across all IPRs in the cell pools, in the absence or presence of a ssODN donor, and with or without NU7441 treatment (n = 2-8). Red: +1 insertion (NHEJ); blue: -7 deletion (MMEJ); green: +2 insertion due to SSTR; black: other indels.

## Panel 7B

```{r Fig7B pathway contribution per barcode with NU7441, fig.height=8, fig.width=14}
pathway_distribution <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios))

barcode_order <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>%
  dplyr::select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("other_indels", "MMEJ", "SSTR", "NHEJ"))

trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  distinct(efficiency, barcode)  %>%
  arrange(dplyr::desc(efficiency)) %>% ggplot(.,
       aes(factor(barcode,levels=barcode_order$barcode), efficiency, width=1)) +
  geom_bar(stat = "identity") +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("TIF") +
  xlab("IPR")

ggplot(pathway_distribution,
       aes(barcode, ratios, width=.8)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  xlab("IPR") +
  scale_fill_manual("legend", values = colores) +
  ggtitle("Pathway contribution per barcode")
```

### Figure legend
(B) Pearson’s correlations of the local intensities of 24 chromatin features versus the relative activity of each pathway (n = 2-3). Correlations with p > 0.001 not shown. Colors as in A.


## Panel 7C

```{r Fig7C SSTR ratio per heterochomatin domain, fig.height=6, fig.width=10}
tib.dt = data.table(trip_tib)

domain_levels = c('late_replicating', 'LAD', 'H3K9me2', 'H3K27me3')

domains = paste0(domain_levels, '_domain')

indel.dt = unique(tib.dt[exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64",  
                         c('exp_pool', 'barcode', 'SSTR', 'SSTR_MMEJSSTR', 'NHEJ', 'MMEJ', 'efficiency', 'SSTR_freq', 'MMEJ_freq','NHEJ_freq',
                           domains), with=F])

indel.dt = indel.dt %>% mutate(NHEJ_eff = NHEJ/efficiency, MMEJ_eff = MMEJ/efficiency, SSTR_eff = SSTR/efficiency)

indel.dt[indel.dt=='iDomain'] <- 0
indel.dt[indel.dt=='Domain'] <- 1

indel.dt[is.na(LAD_domain), LAD_domain:=0]

indel.dt[, (domains) := lapply(.SD, as.numeric), .SDcols = domains]

colnames(indel.dt) = gsub('_domain', '', colnames(indel.dt))

opt_df = get_combinations(domain_levels)


# Ratio of SSTR vs all indels

# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'efficiency', min_count = 25, y_plabel=c(0, 1.1))

plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'SSTR', min_count = 20, y_plabel=c(0.20, 0.26))
# 
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'SSTR_freq', min_count = 25, y_plabel=c(0, 0.5))
# 
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'SSTR_eff', min_count = 25, y_plabel=c(0, 1))
# 
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'MMEJ', min_count = 20, y_plabel=c(.3, 0.4))
# 4
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'MMEJ_freq', min_count = 25, y_plabel=c(0, 0.6))
# 
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'MMEJ_eff', min_count = 25, y_plabel=c(0, 1))
# 
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'NHEJ', min_count = 20, y_plabel=c(.8, .9))
# 
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
#                    domain_levels, 'NHEJ_freq', min_count = 25, y_plabel=c(0, 1))
# 
# plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   # domain_levels, 'NHEJ_eff', min_count = 25, y_plabel=c(0, 1))

```



### Figure legend
(C) Proportion of DSBs repaired by SSTR in the presence of the ssODN donor, for each IPR, split according to the heterochromatin features present as indicated by black dots in the scheme below the graph. Boxed numbers indicate the number of IPRs in each group; only groups with >20 IPRs are shown; see Figure S8I for all groups. Asterisks mark p-values according to the Wilcoxon test, compared to euchromatin IPRs (most left column): * p < 0.05, ** p < 0.01, *** p < 0.001, **** p < 0.0001. 


## Panel 7D

```{r Fig7D MMEJ ratio per heterochomatin domain, fig.height=6, fig.width=10}

plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'MMEJ', min_count = 20, y_plabel=c(.3, 0.4))

```


### Figure legend
(D) Same as in C, but showing the MMEJ:SSTR balance. See also Figure S8J.

```{r Fig7D NHEJ ratio per heterochomatin domain, fig.height=6, fig.width=10}
 
plot_comb_grid(indel.dt[exp_pool=='mean_RSTP2_2000_LBR2_DMSO_ssODN_64'], opt_df,
                   domain_levels, 'NHEJ', min_count = 20, y_plabel=c(.8, .9))

```


# Supplementary Figure 7
## Panel S7A
Schematic of the ssODN
### Figure legend
(A) Schematic of the strategy to probe SSTR simultaneously with NHEJ and MMEJ. Prior to Cas9 activation, the ssODN is co-transfected with a plasmid that encodes the LBR2 sgRNA. The ssODN (black bar) covers the reporter sequence but not the barcode, and contains a 2 bp insertion (green) at the PAM. 


## Panel S7B
```{r FigS7B indel distribution ssODN, fig.height=4, fig.width=7}
SSTR_indels = filter(trip_tib_mut_2000, exp_pool %in% c("mean_RSTP2_2000_LBR2_DMSO_ssODN_64", "mean_RSTP2_2000_LBR2_NU7441_ssODN_64"))

ggplot(SSTR_indels,
       aes(mutation, freq, color = color)) +
  geom_quasirandom() +
  theme_bw(base_size = 16) +
  theme(legend.position = c(0.2, 0.6)) +
  scale_color_manual("legend", values = colory) +
  facet_grid(. ~ drug)

SSTR_indels %>%
  group_by(mutation, drug) %>%
  dplyr::summarise(median = median(freq)) %>%
  arrange(dplyr::desc(median))

SSTR_indels %>%
  group_by(drug) %>%
  distinct(barcode, drug) %>% 
  dplyr::summarise(n_barcodes = n())

trip_tib_mut %>%
  distinct(exp_pool, nexp)

```

### Figure legend
(B) Indel frequencies generated by NHEJ, MMEJ and SSTR in 965 IPRs in the two cell pools, 64 hours after Cas9 activation (average of two replicate experiments). Left panel with DMSO, right panel with 1mM NU7441. N = 2-3. 

## Panel S7C
```{r FigS7C scatter NHEJ with and without ssODN, fig.height=5, fig.width=5}
formula <- y ~ x

trip_lbr2_ssODN_NHEJ <- trip_tib %>% dplyr::select(barcode, ssODN, NHEJ, late_replicating, drug) %>% spread(ssODN, NHEJ)

ggplot(trip_lbr2_ssODN_NHEJ, aes(`-`, ssODN, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_abline(slope = 1)  +
  geom_smooth(method = glm) +
  xlim(0,.8) + ylim(0,.8)  +
  labs(title = "NHEJ - vs NHEJ + ssODN in NU7441 condition",
       x = "NHEJ -",
       y = "NHEJ ssODN") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))
```

### Figure legend
(C-D) Comparison of +1 (NHEJ; panel C) and -7 (MMEJ, panel D) indel frequencies in all IPRs in cell pools in the presence (+) or absence (-) of the ssODN. Black line: diagonal. 


## Panel S7D
```{r FigS7D scatters MMEJ ssODN vs MMEJ - DMSO & NU7441, fig.height=5, fig.width=5}
trip_lbr2_ssODN_MMEJ <- trip_tib %>% dplyr::select(barcode, ssODN, MMEJ, late_replicating, drug) %>% spread(ssODN, MMEJ)

ggplot(trip_lbr2_ssODN_MMEJ, aes(`-`, ssODN, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_abline(slope = 1)  +
  geom_smooth(method = glm) +
  xlim(0,.8) + ylim(0,.8)  +
  labs(title = "MMEJ - vs MMEJ + ssODN in NU7441 condition",
       x = "MMEJ -",
       y = "MMEJ ssODN") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))
```

###  Figure legend



```{r FigS7D scatters efficiency ssODN vs efficiency - DMSO & NU7441, fig.height=5, fig.width=5}
trip_lbr2_ssODN_efficiency <- trip_tib %>% dplyr::select(barcode, ssODN, efficiency, late_replicating, drug) %>% spread(ssODN, efficiency)

ggplot(trip_lbr2_ssODN_efficiency, aes(`-`, ssODN, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_abline(slope = 1)  +
  geom_smooth(method = glm) +
  xlim(0,1) + ylim(0,1)  +
  labs(title = "efficiency - vs efficiency + ssODN in NU7441 condition",
       x = "efficiency -",
       y = "efficiency ssODN") +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "top") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))
```


## Panel S7E

```{r FigS7E MMEJ vs SSTR with DMSO, fig.height=5, fig.width=5}
trip_tib %>%
  filter(ssODN == "ssODN") %>%
  ggplot(., aes(MMEJ, SSTR, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  xlim(0,.5) + ylim(0,.5)   +
  labs(title = "MMEJ vs SSTR in DMSO & NU7441 condition",
       x = "MMEJ",
       y = "SSTR")  +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "bottom") +
  theme(legend.position = c(0.8, 0.2)) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))
```

## Panel S7F
(E) Comparison of MMEJ and SSTR frequencies across all IPRs in cell pools treated with the ssODN. 

```{r FigS7F MMEJ vs SSTR with NU7441, fig.height=5, fig.width=5}
trip_tib %>%
  filter(ssODN == "ssODN") %>%
  ggplot(., aes(efficiency, SSTR, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  labs(title = "efficiency vs SSTR in DMSO & NU7441 condition",
       x = "efficiency",
       y = "SSTR")  +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "bottom") + 
  theme(legend.position = c(0.4, 0.4)) +
  xlim(0,1) + ylim(0,1) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))


trip_tib %>%
  filter(ssODN == "ssODN") %>%
  ggplot(., aes(efficiency, MMEJ, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  labs(title = "efficiency vs MMEJ in DMSO & NU7441 condition",
       x = "efficiency",
       y = "MMEJ")  +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "bottom") + 
  theme(legend.position = c(0.8, 0.4)) +
  xlim(0,1) + ylim(0,1) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))

trip_tib %>%
  filter(ssODN == "ssODN") %>%
  ggplot(., aes(efficiency, NHEJ, color = drug)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  labs(title = "efficiency vs NHEJ in DMSO & NU7441 condition",
       x = "efficiency",
       y = "NHEJ")  +
  stat_cor(method = "pearson", label.x.npc = "left", label.y.npc = "bottom") + 
  theme(legend.position = c(0.8, 0.2)) +
  xlim(0,1) + ylim(0,1) +
  scale_color_manual(values = c("DMSO" = "black", "NU7441" = "gray"))
```

###  Figure legend
(F) Same as (E) but in the presence of NU7441. 


```{r FigS7E SSTR in DMSO vs NU7441, fig.height=5, fig.width=5}
trip_tib %>%
  filter(ssODN == "ssODN") %>%
  dplyr::select(barcode, SSTR, drug) %>%
  tidyr::spread(drug, SSTR) %>%
  ggplot(., aes(DMSO, NU7441)) +
  geom_point() +
  theme_bw(base_size = 16) +
  geom_smooth(method = glm) +
  labs(title = "SSTR in DMSO vs NU7441",
       x = "DMSO",
       y = "NU7441") +
  stat_cor(method = "pearson",label.x.npc = "left", label.y.npc = "top")
```


```{r FigS7B pathway contribution per barcode with DMSO, fig.height=8, fig.width=14}
pathway_distribution <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios))

barcode_order <- trip_tib %>%
  filter(exp_pool == "mean_RSTP2_2000_LBR2_DMSO_ssODN_64") %>%
  gather(NHEJ, MMEJ, SSTR, other_indels,efficiency,
         key = "pathway", value = 'ratios') %>%
    distinct(barcode, pathway, ratios) %>%
  group_by(barcode) %>%
  dplyr::mutate(ratios = ratios/sum(ratios)) %>%
  filter(pathway == "efficiency")  %>%
  arrange(dplyr::desc(ratios)) %>%
  dplyr::select(barcode)

pathway_distribution$barcode <- factor(pathway_distribution$barcode, levels=barcode_order$barcode)
pathway_distribution$pathway <- factor(pathway_distribution$pathway, levels=c("SSTR", "MMEJ", "other_indels",  "NHEJ"))

ggplot(pathway_distribution,
       aes(barcode, ratios, width=1)) +
  geom_bar(stat = "identity", aes(fill = pathway)) +
  theme_bw(base_size = 16) +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  ylab("Pathway Contribution") +
  xlab("IPR") +
  scale_fill_manual("legend", values = colores) +
  ggtitle("Pathway contribution per barcode") 
```


# Bibliography
```{r citations}
cite_packages()
```


# Session Info
```{r}

paste("Run time: ",format(Sys.time()-StartTime))
getwd()
date()
sessionInfo()
```
